### 2020-03-02 README.Rmd
Initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# archyconfgender

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh///master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: </archive/master.zip>

Or you can install this compendium as an R package, archyconfgender, from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
remotes::install_github("/")
```

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2020-03-02 analysis/paper/paper.Rmd
Add Intro

---
title: "Title Goes Here"
author:
  - Myfirstname Mylastname:
      email: fl@oneplace.edu
      institute: [UofO]
      correspondence: true
  - Yourfirstname Yourlastname:
      email: fl@another.edu
      institute: [UofA]
      correspondence: false
institute:
  - UofO: University of One Place
  - UofA: University of Another Place
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

# Introduction

Here is a citation [@Marwick2017]

# Background

Gender underrepresentation and inequality has long existed, from underrepresentation in children’s picture books to academic publications (Hamilton et al. 2006; Tushingham et al. 2017). When talking about inequality in academic fields, we usually refer to the disproportionality between men and women in the field, including archaeology. Women are often underrepresented due to the lack of support women receive in academic fields (Xu 2008). When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field (Bardolph et al. 2016). As we can see in the authorship of the main authors from the Journal of Field Archaeology, 78% of them are men (Heath-Stout 2020). When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, Tushingham et al. (2017) report a trend of women increasingly maintaining their society membership and the gender gap is almost gone. In this study we investigate gender ratios of presenters at major archaeological conferences. We extend the work of Tushingham et al. (2017) with a study of gender ratios of presenters at meetings of the Society of American Archaeology (SAA, 2016-2019), the European Association of Archaeologists (EAA, 2018) and the Computer Applications and Quantitative Methods in Archaeology (CAA, 2017)  We explore how gender ratios vary over time and between these conferences. We examine the influence of the gender of the presenter on the topics that they present on.

# Methods

# Results

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r demo-plot, fig.cap="A plot of random numbers"}
plot(rnorm(10))
```

Figure \@ref(fig:demo-plot) shows how we can have a caption and cross-reference for a plot

```{r demo-inline-code}
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->
\newpage

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2020-03-02 analysis/paper/paper.Rmd
move heading
Name

@@
 author:
-  - Myfirstname Mylastname:
+  - Yichun Chen
       email: fl@oneplace.edu

### 2020-03-02 analysis/paper/paper.Rmd
firstname

@@
       correspondence: true
-  - Yourfirstname Yourlastname:
+  - Ben Yourlastname:
       email: fl@another.edu
@@
 
-# Background
-
 Gender underrepresentation and inequality has long existed, from underrepresentation in children’s picture books to academic publications (Hamilton et al.
@@
 We examine the influence of the gender of the presenter on the topics that they present on.
+
+# Background
 

### 2020-03-02 analysis/paper/paper.Rmd
Last Name

@@
       correspondence: true
-  - Ben Yourlastname:
+  - Yourfirstname Marwick
       email: fl@another.edu

### 2020-03-02 analysis/paper/paper.Rmd
Fix Name

--- 2020-03-02 analysis/paper/paper.Rmd
@@
 author:
-  - Yichun Chen
+  - Yichun Chen:
       email: fl@oneplace.edu
@@
       correspondence: true
-  - Ben Yourlastname:
+  - Ben Marwick:
       email: fl@another.edu

### 2020-03-09 analysis/paper/paper.Rmd
Putting methods

@@
   - Yichun Chen:
-      email: fl@oneplace.edu
+      email: yichun33@uw.edu
       institute: [UofO]
@@
 # Methods
-
+We collected lists of the people that presented research at recent SAA (2016 with 905 presenters, 2017 with 992 presenters, 2018 with 3512 presenters, 2019 with 1014 presenters) EAA (2018 with 2928 presenters) and CAA (2018 with 358 presenters) conferences.
+We obtained publically available programs of these conferences and also tidied the data into a rectangular form with variables including the primary author’s first name, last name and the title of their presentation.
+We used R and the gender package (Mullen 2018; Mihaljević 2019; Belvis et al.
+2015) to predict the gender of all first-named presenters in our sample.
+We used the gender package to identify a first name as male or female, based on a probability scale computed from government records of name-gender data, specifically US Social Security Administration (SSA) baby name data [citation].
+This method has the advantage of speed and ease of automation, but also some substantial limitations.
+We are only able to infer gender into binary male/female genders and assign probabilities for the first names into these categories.
+This has the unfortunate result of excluding other genders from the results, and we encourage conference organisers to collect gender data directly from participants to improve the representation of minority genders in future studies.
+The package often fails to classify non-English names, as the SSA data the package uses are mostly in English names.
+This means that people with non-English names are underrepresented in our results.
+To remove names where gender is ambiguous, which we define as a similar proportion of males and females having that name in the SSA data, we excluded names where the difference in proportions was less than 0.47, a figure we obtained by exploratory data analysis.
+It is important to note that the gender designations here are not self-identified by the presenters, but are computed probabilistically.
+Better quality data will come if presenters self-identify their gender, but we have no alternative because currently these data are not available.
+Our hope is that this work, despite its many limitations, will stimulate the collection of more reliable, justifiable, and useful data by conference organisers.
 # Results
@@
 The current Git commit details are:
-
+u
 ```{r}

### 2020-03-09 analysis/paper/paper.Rmd
Topic Model Method

@@
 Our hope is that this work, despite its many limitations, will stimulate the collection of more reliable, justifiable, and useful data by conference organisers.
+We are also interested in whether there is a  variation in topic choices between the two biological genders in archaeological conferences.
+To do this we have used the process of topic modeling, identifying topics by key words within the topics.
+First, we have to determine the number of topic categories to have and in order to do this we have relied on Latent Dirichlet allocation (LDA).
+LDA finds optimal number of topics then within each topics, keywords are determined then given probability distribution for each keywords.
+We are only interested in keywords that are meaningful in a topic, therefore we have excluded words that are common in the field of archaeology, for example, the, a, el, al, archaeology...etc.
 # Results

### 2020-03-09 analysis/paper/paper.Rmd
little bit of editing

@@
 
-Gender underrepresentation and inequality has long existed, from underrepresentation in children’s picture books to academic publications (Hamilton et al.
+Gender under-representation and inequality has long existed, from under-representation in children’s picture books to academic publications (Hamilton et al.
 2006; Tushingham et al.
@@
 # Methods
-We collected lists of the people that presented research at recent SAA (2016 with 905 presenters, 2017 with 992 presenters, 2018 with 3512 presenters, 2019 with 1014 presenters) EAA (2018 with 2928 presenters) and CAA (2018 with 358 presenters) conferences.
-We obtained publically available programs of these conferences and also tidied the data into a rectangular form with variables including the primary author’s first name, last name and the title of their presentation.
+
+```{r}
+# CTRL + OPTION + i to insert a code block
+
+saa_2016_n_presenters <- 905
+```
+
+We collected lists of the people that presented research at recent SAA (2016 with `r saa_2016_n_presenters` presenters, 2017 with 992 presenters, 2018 with 3512 presenters, 2019 with 1014 presenters) EAA (2018 with 2928 presenters) and CAA (2018 with 358 presenters) conferences.
+We obtained publicly available programs of these conferences and also tidied the data into a rectangular form with variables including the primary author’s first name, last name and the title of their presentation.
 We used R and the gender package (Mullen 2018; Mihaljević 2019; Belvis et al.
@@
 We are only able to infer gender into binary male/female genders and assign probabilities for the first names into these categories.
-This has the unfortunate result of excluding other genders from the results, and we encourage conference organisers to collect gender data directly from participants to improve the representation of minority genders in future studies.
+This has the unfortunate result of excluding other genders from the results, and we encourage conference organizers to collect gender data directly from participants to improve the representation of minority genders in future studies.
 The package often fails to classify non-English names, as the SSA data the package uses are mostly in English names.
@@
 Our hope is that this work, despite its many limitations, will stimulate the collection of more reliable, justifiable, and useful data by conference organisers.
+
 We are also interested in whether there is a variation in topic choices between the two biological genders in archaeological conferences.
-To do this we have used the process of topic modeling, identifying topics by key words within the topics.
+To do this we have used the process of topic modeling, identifying topics by keywords within the topics.
 First, we have to determine the number of topic categories to have and in order to do this we have relied on Latent Dirichlet allocation (LDA).
@@
 We are only interested in keywords that are meaningful in a topic, therefore we have excluded words that are common in the field of archaeology, for example, the, a, el, al, archaeology...etc.
+
 # Results

### 2020-03-23 analysis/paper/paper.Rmd
Revision

@@
 We are also interested in whether there is a variation in topic choices between the two biological genders in archaeological conferences. 
-To do this we have used the process of topic modeling, identifying topics by keywords within the topics.
-First, we have to determine the number of topic categories to have and in order to do this we have relied on Latent Dirichlet allocation (LDA).
-LDA finds optimal number of topics then within each topics, keywords are determined then given probability distribution for each keywords.
-We are only interested in keywords that are meaningful in a topic, therefore we have excluded words that are common in the field of archaeology, for example, the, a, el, al, archaeology...etc.
+To do this we have used the process of topic modeling, which is used to find natural groups of words when we do not have a specific criteria or simply unsure what we are looking for.
+There are many ways to accomplish topic modeling, we have used the method of Latent Dirichlet allocation (LDA).
+This method is beneficial in our case because LDA does not treat one topic, which consists of many words, as seperate from another topic.
+Before we can start using LDA, we have to first find the number of groups that we need to predetermine.
+To do this we have used the package ldatuning, which allows us to use multiple methods at once to find the optimal number of groups that we should use for LDA processing.
+With the number of groups figured out, we then create a topic model based on the number of topics determined, from this we can then assign the words a probability based on its per-topic-per-word probability, which we call beta.
+After this we filter out only the top 10 terms within each group to keep so that we have the keywords that occur the most by probability.
+From this we then have top terms from the topics of the conferences.
+We are only interested in keywords that are meaningful in a topic, therefore we have excluded words that are common in the field of archaeology or simply meaningless, for example, the, a, el, al, archaeology...etc.
+After figuring out the top terms, we can combine them with our earlier exploration of gender to see if one gender seems to favor another, for example if female or male favor topics about 'rock'.
 

### 2020-03-23 analysis/paper/paper.Rmd
Adding 1 sentence

@@
 To do this we have used the package ldatuning, which allows us to use multiple methods at once to find the optimal number of groups that we should use for LDA processing.
-With the number of groups figured out, we then create a topic model based on the number of topics determined, from this we can then assign the words a probability based on its per-topic-per-word probability, which we call beta.
+With the number of groups figured out, we then create a topic model based on the number of topics determined, this will put words into groups that most frequently occur together, from this we can then assign the words a probability based on its per-topic-per-word probability, which we call beta.
 After this we filter out only the top 10 terms within each group to keep so that we have the keywords that occur the most by probability.

### 2020-03-23 analysis/paper/paper.Rmd
Added some code to simplify

@@
 
-# Methods
-
 ```{r}
 # CTRL + OPTION + i to insert a code block
-
-saa_2016_n_presenters <- 905
+SAA16 <- readxl::read_excel("SAA2016.xlsx")
+SAA17 <- readxl::read_excel("SAA2017.xlsx")
+SAA18 <- readxl::read_excel("SAA2018-finalprogram.xlsx")
+SAA19 <- readxl::read_excel("SAA2019.xlsx")
+EAA18 <- readxl::read_excel("2018 EAA.xlsx")
+CAA18 <- readxl::read_excel("2018 CAA.xlsx")
+SAA18_top_term_plot <- readRDS("SAA18 top term plot")
+SAA16_row <- nrow(SAA16)
+SAA17_row <- nrow(SAA17)
+SAA18_row <- nrow(SAA18)
+SAA19_row <- nrow(SAA19)
+EAA18_row <- nrow(EAA18)
+CAA18_row <- nrow(CAA18)
 ```
-
-We collected lists of the people that presented research at recent SAA (2016 with `r saa_2016_n_presenters` presenters, 2017 with 992 presenters, 2018 with 3512 presenters, 2019 with 1014 presenters) EAA (2018 with 2928 presenters) and CAA (2018 with 358 presenters) conferences.
+# Methods
+We collected lists of the people that presented research at recent SAA (2016 with `r SAA16_row` presenters, 2017 with `r SAA17_row` presenters, 2018 with `r SAA18_row` presenters, 2019 with `r SAA19_row` presenters) EAA (2018 with `r EAA18_row` presenters) and CAA (2018 with `r CAA18_row` presenters) conferences.
 We obtained publicly available programs of these conferences and also tidied the data into a rectangular form with variables including the primary author’s first name, last name and the title of their presentation.
@@
 After this we filter out only the top 10 terms within each group to keep so that we have the keywords that occur the most by probability.
-From this we then have top terms from the topics of the conferences.
+Below is an example of a plot for the 2018 SAA that shows the predetermined number of groups each with top 10 of the keywords in each group.
+`r SAA18_top_term_plot` From this we then have top terms from the topics of the conferences.
 We are only interested in keywords that are meaningful in a topic, therefore we have excluded words that are common in the field of archaeology or simply meaningless, for example, the, a, el, al, archaeology...etc.

### 2020-03-23 analysis/paper/paper.Rmd
make the code work, the document render, and tidy up

--- 2020-03-23 analysis/paper/paper.Rmd
@@
 
+```{r}
+# CTRL + OPTION + i to insert a code block
+library(here)
+SAA16 <- readxl::read_excel(here("analysis/data/SAA2016.xlsx"))
+SAA17 <- readxl::read_excel(here("analysis/data/SAA2017.xlsx"))
+SAA18 <- readxl::read_excel(here("analysis/data/SAA2018-finalprogram.xlsx"))
+SAA19 <- readxl::read_excel(here("analysis/data/SAA2019.xlsx"))
+EAA18 <- readxl::read_excel(here("analysis/data/2018 EAA.xlsx"))
+CAA18 <- readxl::read_excel(here("analysis/data/2018 CAA.xlsx"))
+SAA18_top_term_plot <- readRDS(here("analysis/data/SAA18 top term plot"))
+
+SAA16_row <- nrow(SAA16)
+SAA17_row <- nrow(SAA17)
+SAA18_row <- nrow(SAA18)
+SAA19_row <- nrow(SAA19)
+EAA18_row <- nrow(EAA18)
+CAA18_row <- nrow(CAA18)
+
+```
+
 # Methods
 
-```{r}
-# CTRL + OPTION + i to insert a code block
-
-saa_2016_n_presenters <- 905
-```
-
-We collected lists of the people that presented research at recent SAA (2016 with `r saa_2016_n_presenters` presenters, 2017 with 992 presenters, 2018 with 3512 presenters, 2019 with 1014 presenters) EAA (2018 with 2928 presenters) and CAA (2018 with 358 presenters) conferences.
+We collected lists of the people that presented research at recent SAA (2016 with `r SAA16_row` presenters, 2017 with `r SAA17_row` presenters, 2018 with `r SAA18_row` presenters, 2019 with `r SAA19_row` presenters) EAA (2018 with `r EAA18_row` presenters) and CAA (2018 with `r CAA18_row` presenters) conferences.
 We obtained publicly available programs of these conferences and also tidied the data into a rectangular form with variables including the primary author’s first name, last name and the title of their presentation.
@@
 We are also interested in whether there is a variation in topic choices between the two biological genders in archaeological conferences.
+
 To do this we have used the process of topic modeling, which is used to find natural groups of words when we do not have a specific criteria or simply unsure what we are looking for.
@@
 After this we filter out only the top 10 terms within each group to keep so that we have the keywords that occur the most by probability.
+Below is an example of a plot for the 2018 SAA that shows the predetermined number of groups each with top 10 of the keywords in each group.
 From this we then have top terms from the topics of the conferences.
@@
 After figuring out the top terms, we can combine them with our earlier exploration of gender to see if one gender seems to favor another, for example if female or male favor topics about 'rock'.
+
+```{r}
+SAA18_top_term_plot
+```
 
@@
 # Note the path that we need to use to access our data files when rendering this document
-my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
+# my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
 ```

### 2020-03-23 README.Rmd
add binder and actions

@@
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh///master?urlpath=rstudio)
+[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/yichun33/archyconfgender/master?urlpath=rstudio)
+[![R build status](https://github.com/yichun33/archyconfgender/workflows/R-CMD-check/badge.svg)](https://github.com/yichun33/archyconfgender/actions)
 

### 2020-03-23 analysis/paper/paper.Rmd
a bit of editing

@@
 There are many ways to accomplish topic modeling, we have used the method of Latent Dirichlet allocation (LDA).
-This method is beneficial in our case because LDA does not treat one topic, which consists of many words, as seperate from another topic.
-Before we can start using LDA, we have to first find the number of groups that we need to predetermine.
-To do this we have used the package ldatuning, which allows us to use multiple methods at once to find the optimal number of groups that we should use for LDA processing.
-With the number of groups figured out, we then create a topic model based on the number of topics determined, this will put words into groups that most frequently occur together, from this we can then assign the words a probability based on its per-topic-per-word probability, which we call beta.
-After this we filter out only the top 10 terms within each group to keep so that we have the keywords that occur the most by probability.
-Below is an example of a plot for the 2018 SAA that shows the predetermined number of groups each with top 10 of the keywords in each group.
-From this we then have top terms from the topics of the conferences.
+This method is beneficial in our case because LDA does not treat one topic, which consists of many words, as separate from another topic.
+Before we can start using LDA, we have to first decide on the number of topics in our collection of texts.
+To do this we have used the R package ldatuning (add citation), which allows us to use multiple methods to compute an optimal number of groups for LDA processing.
+We then create a topic model for all our texts, this will assign all words in every document into topics, and assign each word a probability based on its per-topic-per-word probability, which we call beta.
+After this we keep only the top 10 words within each topic as keywords to represent each topic.
+Below is an example of a plot for the 2018 SAA that shows the predetermined number of groups each with top ten keywords in each group.
+From this we then have top terms from the topics of all conferences.
 We are only interested in keywords that are meaningful in a topic, therefore we have excluded words that are common in the field of archaeology or simply meaningless, for example, the, a, el, al, archaeology...etc.

### 2020-04-10 analysis/paper/function.R
editing the text

@@ -1,0 +1,45 @@
+
+
+# function to compute topic model for a conf
+
+do_the_topic_model <- function(x) {
+
+  d <- readxl::read_excel(here::here(x))
+
+  library(tidyverse)
+  library(tidytext)
+
+  d_word <- d %>%
+    unnest_tokens(word, AbstractTitle)
+
+  word_counts <- d_word %>%
+    anti_join(stop_words) %>%
+    filter(!word %in% c("de", "la", "el", "los", "user",
+                        "archaeology", "archaeological"
+                        ,"site", "archaic", "sites")) %>%
+    filter(!str_detect(word, "\\d")) %>%
+    filter(str_length(word) >= 3) %>%
+    filter(!str_detect(word, "\\.|'|\\<U")) %>%
+    unite(unique_id, SessionNumber, LastName, FirstName) %>%
+    count(unique_id, word, sort = TRUE) %>%
+    ungroup()
+
+  unique_id_dtm <- word_counts %>%
+    cast_dtm(unique_id, word, n)
+
+  library(topicmodels)
+  unique_id_lda <- LDA(unique_id_dtm, k = 21, control = list(seed = 1234))
+
+  top_terms <- unique_id_lda %>%
+    group_by(topic) %>%
+    top_n(10, beta) %>%
+    ungroup() %>%
+    arrange(topic, -beta)
+
+  unique_id_gamma <- tidy(unique_id_lda, matrix = "gamma")
+
+  unique_id_gamma <- unique_id_gamma %>%
+    separate(document, c("SessionNumber", "LastName", "FirstName"), sep = "_", convert = TRUE)
+}
+
+do_the_topic_model("")

### 2020-04-10 analysis/paper/function.R
making a fn for topic modelling

@@ -2,9 +2,11 @@
 
 # function to compute topic model for a conf
 
+# x is the data frame from the spreadsheet
+
 do_the_topic_model <- function(x) {
 
-  d <- readxl::read_excel(here::here(x))
+  d <- x
 
   library(tidyverse)
   library(tidytext)
@@ -20,7 +22,7 @@ do_the_topic_model <- function(x) {
     filter(!str_detect(word, "\\d")) %>%
     filter(str_length(word) >= 3) %>%
     filter(!str_detect(word, "\\.|'|\\<U")) %>%
-    unite(unique_id, SessionNumber, LastName, FirstName) %>%
+    unite(unique_id, session, FirstName) %>%
     count(unique_id, word, sort = TRUE) %>%
     ungroup()
 
@@ -39,7 +41,7 @@ do_the_topic_model <- function(x) {
   unique_id_gamma <- tidy(unique_id_lda, matrix = "gamma")
 
   unique_id_gamma <- unique_id_gamma %>%
-    separate(document, c("SessionNumber", "LastName", "FirstName"), sep = "_", convert = TRUE)
+    separate(document, c("session", "FirstName"), sep = "_", convert = TRUE)
 }
 
-do_the_topic_model("")
+CAA18_tm <- do_the_topic_model(CAA18)

### 2020-04-10 analysis/paper/function.R
working on fn

@@ -30,9 +30,12 @@ do_the_topic_model <- function(x) {
     cast_dtm(unique_id, word, n)
 
   library(topicmodels)
-  unique_id_lda <- LDA(unique_id_dtm, k = 21, control = list(seed = 1234))
+  unique_id_lda <- LDA(unique_id_dtm,
+                       k = 21,
+                       control = list(seed = 1234))
 
-  top_terms <- unique_id_lda %>%
+  top_terms <-  tidy(unique_id_lda,
+                     matrix = "beta")  %>%
     group_by(topic) %>%
     top_n(10, beta) %>%
     ungroup() %>%
@@ -41,7 +44,11 @@ do_the_topic_model <- function(x) {
   unique_id_gamma <- tidy(unique_id_lda, matrix = "gamma")
 
   unique_id_gamma <- unique_id_gamma %>%
-    separate(document, c("session", "FirstName"), sep = "_", convert = TRUE)
+    separate(document, c("session", "FirstName"),
+             sep = "_",
+             convert = TRUE)
+
+  return()
 }
 
 CAA18_tm <- do_the_topic_model(CAA18)

### 2020-04-10 analysis/paper/paper.Rmd
prepare homework

@@
 library(here)
-SAA16 <- readxl::read_excel(here("analysis/data/SAA2016.xlsx"))
-SAA17 <- readxl::read_excel(here("analysis/data/SAA2017.xlsx"))
-SAA18 <- readxl::read_excel(here("analysis/data/SAA2018-finalprogram.xlsx"))
-SAA19 <- readxl::read_excel(here("analysis/data/SAA2019.xlsx"))
-EAA18 <- readxl::read_excel(here("analysis/data/2018 EAA.xlsx"))
-CAA18 <- readxl::read_excel(here("analysis/data/2018 CAA.xlsx"))
-SAA18_top_term_plot <- readRDS(here("analysis/data/SAA18 top term plot"))
+# column names should be:
+# session, FirstName, AbstractTitle
 
-SAA16_row <- nrow(SAA16)
-SAA17_row <- nrow(SAA17)
+SAA18 <- readxl::read_excel(here("analysis/data/raw-data/2018 SAA Title.xlsx"))
+EAA18 <- readxl::read_excel(here("analysis/data/raw-data/2018 EAA.xlsx"))
+CAA18 <- readxl::read_excel(here("analysis/data/raw-data/2018 CAA.xlsx"))
+
 SAA18_row <- nrow(SAA18)      
-SAA19_row <- nrow(SAA19)
 EAA18_row <- nrow(EAA18)      
@@
 
-We collected lists of the people that presented research at recent SAA (2016 with `r SAA16_row` presenters, 2017 with `r SAA17_row` presenters, 2018 with `r SAA18_row` presenters, 2019 with `r SAA19_row` presenters) EAA (2018 with `r EAA18_row` presenters) and CAA (2018 with `r CAA18_row` presenters) conferences.
+We collected lists of the people that presented research at recent SAA (2018 with `r SAA18_row` presenters, EAA (2018 with `r EAA18_row` presenters) and CAA (2018 with `r CAA18_row` presenters) conferences.
 We obtained publicly available programs of these conferences and also tidied the data into a rectangular form with variables including the primary author’s first name, last name and the title of their presentation.
@@
 
-We are also interested in whether there is a variation in topic choices between the two biological genders in archaeological conferences.
+We are also interested in whether there is a variation in topic choices between these two genders in archaeological conferences.
+To do this we have used topic modeling, which is a method to automatically find related groups of words that resemble 'themes' or 'topics'.
+We used Latent Dirichlet allocation (LDA) because LDA does not treat one topic, which consists of many words, as separate from another topic.
+That is, a word can be present in many topics but with differnt weightings.
+When using LDA, we have to first decide on the number of topics that the method will identify in our texts.
+We used the R package ldatuning [citation] to compute an optimal number of topics for the LDA method to identify.
+We then create a topic model for all our texts, this will assign all words in every document into our set number of topics, and assign each word a probability based on its per-topic-per-word probability.
+To visualise the topic model we use the top 10 words within each topic as keywords to represent each topic.
+We excluded words that are common in the field of archaeology or have little semantic value, for example, 'the', 'a', 'el', 'al', 'archaeology', etc.
+We generated one topic model per conference.
+Each presentation at each conference is then assigned a vector of topic probabilities.
 
-To do this we have used the process of topic modeling, which is used to find natural groups of words when we do not have a specific criteria or simply unsure what we are looking for.
-There are many ways to accomplish topic modeling, we have used the method of Latent Dirichlet allocation (LDA).
-This method is beneficial in our case because LDA does not treat one topic, which consists of many words, as separate from another topic.
-Before we can start using LDA, we have to first decide on the number of topics in our collection of texts.
-To do this we have used the R package ldatuning (add citation), which allows us to use multiple methods to compute an optimal number of groups for LDA processing.
-We then create a topic model for all our texts, this will assign all words in every document into topics, and assign each word a probability based on its per-topic-per-word probability, which we call beta.
-After this we keep only the top 10 words within each topic as keywords to represent each topic.
-Below is an example of a plot for the 2018 SAA that shows the predetermined number of groups each with top ten keywords in each group.
-From this we then have top terms from the topics of all conferences.
-We are only interested in keywords that are meaningful in a topic, therefore we have excluded words that are common in the field of archaeology or simply meaningless, for example, the, a, el, al, archaeology...etc.
-After figuring out the top terms, we can combine them with our earlier exploration of gender to see if one gender seems to favor another, for example if female or male favor topics about 'rock'.
-
-```{r}
-SAA18_top_term_plot
-```
+To explore the relationship between gender and topics we identified the most prominent topic for each presentation...
 
@@
 
-```{r get-data, eval = FALSE}
-# Note the path that we need to use to access our data files when rendering this document
-# my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
+```{r}
+# make the custom functions available here in this Rmd file
+source(here::here("analysis/paper/functions-for-gender.R"))
+source(here::here("analysis/paper/functions-for-topic-models.R"))
 ```
 
-```{r demo-plot, fig.cap="A plot of random numbers"}
-plot(rnorm(10))
+```{r}
+# let's run the topic model fn for each conf
+CAA18_tm <- do_the_topic_model(CAA18)
+xxx_tm <- do_the_topic_model(xxx)
+xxx_tm <- do_the_topic_model(xxx)
+
+# Let's save the output for each one
+saveRDS(CAA18_tm, here::here("analysis/data/derived-data/CAA18_topic_model.rds"))
+saveRDS(xxx_tm, here::here("analysis/data/derived-data/xxx_topic_model.rds"))
+saveRDS(xxx_tm, here::here("analysis/data/derived-data/xxx_topic_model.rds"))
 ```

### 2020-04-10 analysis/paper/functions-for-topic-models.R

@@ -4,7 +4,7 @@
 
 # x is the data frame from the spreadsheet
 
-do_the_topic_model <- function(x) {
+do_the_topic_model <- function(x, k = 21) {
 
   d <- x
 
@@ -31,7 +31,7 @@ do_the_topic_model <- function(x) {
 
   library(topicmodels)
   unique_id_lda <- LDA(unique_id_dtm,
-                       k = 21,
+                       k = k,
                        control = list(seed = 1234))
 
   top_terms <-  tidy(unique_id_lda,
@@ -48,7 +48,31 @@ do_the_topic_model <- function(x) {
              sep = "_",
              convert = TRUE)
 
-  return()
+  return(list(unique_id_dtm = unique_id_dtm,
+              unique_id_lda = unique_id_lda,
+              top_terms = top_terms,
+              unique_id_gamma = unique_id_gamma))
+
 }
 
-CAA18_tm <- do_the_topic_model(CAA18)
+
+
+# access the different elements with $
+
+# compute the optimum number of topics
+
+compute_optimum_n_topics <- function(dtm){
+  ldatuning::FindTopicsNumber(
+    dtm,
+   topics = seq(from = 2, to = 500, by = 1),
+   metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
+   method = "Gibbs",
+   control = list(seed = 77),
+   mc.cores = 2L,
+   verbose = TRUE
+ )
+}
+
+# CAA18_tm_n_topics <- compute_optimum_n_topics(CAA18_tm$unique_id_dtm)
+# ldatuning::FindTopicsNumber_plot(CAA18_tm_n_topics)
+

### 2020-04-10 analysis/paper/paper.Rmd
Citation

@@
 We collected lists of the people that presented research at recent SAA (2018 with `r SAA18_row` presenters, EAA (2018 with `r EAA18_row` presenters) and CAA (2018 with `r CAA18_row` presenters) conferences.
-We obtained publicly available programs of these conferences and also tidied the data into a rectangular form with variables including the primary author’s first name, last name and the title of their presentation.
-We used R and the gender package (Mullen 2018; Mihaljević 2019; Belvis et al.
+We obtained publicly available programs of these conferences and also tidied the data into a rectangular form with variables including the primary author's first name, last name and the title of their presentation.
+We used R and the gender package (Mullen 2018; Mihaljevic 2019; Belvis et al.
 2015) to predict the gender of all first-named presenters in our sample.
-We used the gender package to identify a first name as male or female, based on a probability scale computed from government records of name-gender data, specifically US Social Security Administration (SSA) baby name data [citation].
+We used the gender package to identify a first name as male or female, based on a probability scale computed from government records of name-gender data, specifically US Social Security Administration (SSA) baby name data (Mullen 2019).
 This method has the advantage of speed and ease of automation, but also some substantial limitations.
@@
 When using LDA, we have to first decide on the number of topics that the method will identify in our texts.
-We used the R package ldatuning [citation] to compute an optimal number of topics for the LDA method to identify.
+We used the R package ldatuning (Nikita 2020) to compute an optimal number of topics for the LDA method to identify.
 We then create a topic model for all our texts, this will assign all words in every document into our set number of topics, and assign each word a probability based on its per-topic-per-word probability.

### 2020-04-17 analysis/paper/paper.Rmd
Citation and do the topic model

@@
 # let's run the topic model fn for each conf
-CAA18_tm <- do_the_topic_model(CAA18)
-xxx_tm <- do_the_topic_model(xxx)
-xxx_tm <- do_the_topic_model(xxx)
+CAA18_tm <- do_the_topic_model(CAA18, 21)
+EAA18_tm <- do_the_topic_model(EAA18, 12)
+SAA18_tm <- do_the_topic_model(SAA18, 21)
 
@@
 saveRDS(CAA18_tm, here::here("analysis/data/derived-data/CAA18_topic_model.rds"))
-saveRDS(xxx_tm, here::here("analysis/data/derived-data/xxx_topic_model.rds"))
-saveRDS(xxx_tm, here::here("analysis/data/derived-data/xxx_topic_model.rds"))
+saveRDS(EAA18_tm, here::here("analysis/data/derived-data/EAA18_topic_model.rds"))
+saveRDS(SAA18_tm, here::here("analysis/data/derived-data/SAA18_topic_model.rds"))
 ```

### 2020-04-17 analysis/paper/functions-for-topic-models.R

@@ -4,7 +4,7 @@
 
 # x is the data frame from the spreadsheet
 
-do_the_topic_model <- function(x, k = 21) {
+do_the_topic_model <- function(x, k) {
 
   d <- x
 

### 2020-04-17 analysis/paper/paper.Rmd
GM

@@
 saveRDS(SAA18_tm, here::here("analysis/data/derived-data/SAA18_topic_model.rds"))
+
+# gender model
+CAA18_gm <- get_the_gender_data(CAA18)
+saveRDS(CAA18_gm, here::here("analysis/data/derived-data/CAA18_gender_model.rds"))
+SAA18_gm <- get_the_gender_data(SAA18)
+saveRDS(SAA18_gm, here::here("analysis/data/derived-data/SAA18_gender_model.rds"))
 ```

### 2020-04-17 analysis/paper/functions-for-joining-table.R

@@ -1,0 +1,140 @@
+
+
+# functions for joining table
+
+# x is dataframe, y is dataframe
+
+joining_table_topic_and_gender <- function(x, y)
+
+library(dplyr)
+library(tidyverse)
+joined_table <-
+  left_join(readRDS("gamma"),
+            readRDS("gender unnest"),
+            by = c("FirstName" = "name",
+                   "SessionNumber" = "SessionNumber",
+                   "LastName" = "LastName"))
+saveRDS(joined_table, "final table")
+
+#Making table similar to Stats Rethinking
+joined_table_wider <-
+joined_table %>%
+  pivot_wider(names_from = "topic",
+              values_from = "gamma")
+
+old_names <- names(joined_table_wider)[5:25]
+new_names <- paste0("topic", old_names)
+renamed_joined_table <-
+joined_table_wider %>%
+  rename_at(vars(old_names), ~new_names) %>%
+  mutate(gender_b = ifelse(gender == "female", 1, 0))
+
+
+joined_table %>%
+  filter(!is.na(gender)) %>%
+  filter(gamma >= 0.1) %>%
+ggplot(aes(x = as.factor(topic), y = gamma)) +
+  geom_boxplot() +
+  facet_wrap(~gender, ncol = 1)
+
+
+joined_table %>%
+  filter(!is.na(gender)) %>%
+  filter(gamma >= 0.1) %>%
+  mutate(new_gamma = ((gamma - mean(gamma)) / sd(gamma))^(1/2)) %>%
+  ggplot(aes(x = as.factor(topic),
+             y = new_gamma,
+             colour = gender)) +
+  geom_boxplot()
+
+
+joined_table %>%
+  filter(!is.na(gender)) %>%
+  filter(gamma >= 0.1) %>%
+  mutate(new_gamma = ((gamma - mean(gamma)) / sd(gamma))^(1/2)) %>%
+  filter(!is.nan(new_gamma)) %>%
+  mutate_at(vars(topic, gender),
+            as.factor) %>%
+  aov(gamma ~ gender * topic, data = . ) %>%
+  summary
+
+d <-
+  joined_table %>%
+  filter(!is.na(gender)) %>%
+ # filter(gamma >= 0.1) %>%
+  mutate(new_gamma = ((gamma - mean(gamma)) / sd(gamma))^(1/2)) %>%
+  filter(!is.nan(new_gamma)) %>%
[…]

### 2020-04-17 analysis/paper/functions-for-joining-table.R
joining table

@@ -2,15 +2,15 @@
 
 # functions for joining table
 
-# x is dataframe, y is dataframe
+# x is a _tm file, y is dataframe
 
 joining_table_topic_and_gender <- function(x, y)
 
 library(dplyr)
 library(tidyverse)
 joined_table <-
-  left_join(readRDS("gamma"),
-            readRDS("gender unnest"),
+  left_join(x$unique_id_gamma,
+            x$genders_unnest,
             by = c("FirstName" = "name",
                    "SessionNumber" = "SessionNumber",
                    "LastName" = "LastName"))

### 2020-05-01 analysis/paper/paper.Rmd
STM

@@
 
+We then tried another process of topic modeling called Structural Topic Model (STM) which is very similar to the LDA we have described above.
+The main advantage for us is that, the STM allows us to include, alongside the normal topic modeling process, other metadata.
+This process "attaches" the metadata to the collection of texts that we will then perform topic modeling process.
+It is important to note that we have used the stm package that also includes the ability to calculate the number of topics that is best suited for STM modeling (Roberts 2019).
+The fact that we are allowed to do this allows us to perform covariate analysis (Roberts et al.
+2013).
+This basically means that we can examine the relationship between the topics and some covariates/variables of interest.
+We can define the genders of males and females and then analyze the prevelance of these two categories in accordance with the topics that the stm package determined (Roberts 2019).
+The analysis is used also with probabilities given to the texts to see if the most occuring of some text seems to be "attached" to one gender over another.
+
 To explore the relationship between gender and topics we identified the most prominent topic for each presentation...

### 2020-05-01 analysis/paper/paper.Rmd
add set seed

@@
 To do this we have used topic modeling, which is a method to automatically find related groups of words that resemble 'themes' or 'topics'.
-We used Latent Dirichlet allocation (LDA) because LDA does not treat one topic, which consists of many words, as separate from another topic.
-That is, a word can be present in many topics but with differnt weightings.
-When using LDA, we have to first decide on the number of topics that the method will identify in our texts.
-We used the R package ldatuning (Nikita 2020) to compute an optimal number of topics for the LDA method to identify.
+This approach does not treat one topic, which consists of many words, as separate from another topic.
+That is, a word can be present in many topics but with differs weightings.
+When generating topic models, we have to first decide on the number of topics that the method will identify in our texts.
 We then create a topic model for all our texts, this will assign all words in every document into our set number of topics, and assign each word a probability based on its per-topic-per-word probability.
@@
 # Results
+```{r}
+number_of_papers_saa_2018 <-  ...
+```
+
+- some basic details ( 1 para)
+-- number of papers that are in our sample for each conference
+-- number of men and women first-named presented, state the gender ratio 
+-- number of people that we could not assign a gender to
+
+The number of ... `r number_of_papers_saa_2018` ...
+
+- the main details
+-- number of topics that are optimum for each conf (1 para)
+-- variance of topics with gender for each conf (1 para)
 

### 2020-05-01 analysis/paper/structural-topic-model.R

@@ -26,13 +26,13 @@ docs <- out$documents
 vocab <- out$vocab
 meta <- out$meta
 
-
 out <-
   prepDocuments(processed$documents,
                 processed$vocab,
                 processed$meta,
                 lower.thresh = 15)
 
+set.seed(33)
 model <-
   stm(
     documents = out$documents,
@@ -41,7 +41,6 @@ model <-
     # watch the output to see how many topics it picks
     prevalence =  ~ gender,
     data = out$meta,
-    max.em.its = 10,
     init.type = "Spectral"
   )
 

### 2020-05-03 analysis/paper/structural-topic-model.R
set seed in stm function

@@ -35,6 +35,7 @@ out <-
 set.seed(33)
 model <-
   stm(
+    seed = 33,
     documents = out$documents,
     vocab = out$vocab,
     K = 0,
@@ -170,6 +171,9 @@ ggplot() +
             colour = "red") +
   xlab("        ← more men  ...  more women →") +
   ylab("Topic")
+
+ggsave(here::here("analysis/figures/stm-covariate-plot.png"),
+       height = 10, width = 6)
 
 
 

### 2020-05-03 analysis/paper/paper.Rmd
we've got plots for all three confs now

@@
 ```{r}
-# CTRL + OPTION + i to insert a code block
 library(here)
@@
 
-SAA18 <- readxl::read_excel(here("analysis/data/raw-data/2018 SAA Title.xlsx"))
-EAA18 <- readxl::read_excel(here("analysis/data/raw-data/2018 EAA.xlsx"))
-CAA18 <- readxl::read_excel(here("analysis/data/raw-data/2018 CAA.xlsx"))
+SAA18 <- readxl::read_excel(here::here("analysis/data/raw-data/2018 SAA Title.xlsx"))
+EAA18 <- readxl::read_excel(here::here("analysis/data/raw-data/2018 EAA.xlsx")) 
+CAA18 <- readxl::read_excel(here::here("analysis/data/raw-data/2018 CAA.xlsx"))
 
@@
 
-```{r}
-# make the custom functions available here in this Rmd file
+```{r, eval=FALSE}
+
+# make the custom function available here in this Rmd file
 source(here::here("analysis/paper/functions-for-gender.R"))
-source(here::here("analysis/paper/functions-for-topic-models.R"))
+
+# compute gender model, take a few minutes...
+SAA18_gm <- get_the_gender_data(SAA18)
+saveRDS(SAA18_gm, here::here("analysis/data/derived-data/SAA18_gender_model.rds")) # ok
+
+EAA18_gm <- get_the_gender_data(EAA18, first_name_appears_first = FALSE)
+saveRDS(EAA18_gm, here::here("analysis/data/derived-data/EAA18_gender_model.rds"))
+
+CAA18_gm <- get_the_gender_data(CAA18)
+saveRDS(CAA18_gm, here::here("analysis/data/derived-data/CAA18_gender_model.rds"))
 ```
 
-```{r}
-# let's run the topic model fn for each conf
-CAA18_tm <- do_the_topic_model(CAA18, 21)
-EAA18_tm <- do_the_topic_model(EAA18, 12)
-SAA18_tm <- do_the_topic_model(SAA18, 21)
+```{r, eval=FALSE}
 
-# Let's save the output for each one
-saveRDS(CAA18_tm, here::here("analysis/data/derived-data/CAA18_topic_model.rds"))
-saveRDS(EAA18_tm, here::here("analysis/data/derived-data/EAA18_topic_model.rds"))
-saveRDS(SAA18_tm, here::here("analysis/data/derived-data/SAA18_topic_model.rds"))
+# make the custom function available here in this Rmd file
+source(here::here("analysis/paper/function-for-structural-topic-model.R"))
 
-# gender model
-CAA18_gm <- get_the_gender_data(CAA18)
-saveRDS(CAA18_gm, here::here("analysis/data/derived-data/CAA18_gender_model.rds"))
-SAA18_gm <- get_the_gender_data(SAA18)
-saveRDS(SAA18_gm, here::here("analysis/data/derived-data/SAA18_gender_model.rds"))
+# compute and save the topic model for each conf, takes quite a long time
+SAA18_stm <- 
+  stm_model_and_plot(conf_paper_titles = SAA18, 
+                     conf_gender_results = SAA18_gm, 
+                     max.em.its = 1000)
+
+saveRDS(SAA18_stm, here::here("analysis/data/derived-data/SAA18_topic_model_and_plot.rds"))
+
+CAA18_stm <-   
+  stm_model_and_plot(conf_paper_titles = CAA18, 
+                     conf_gender_results = CAA18_gm, 
+                     lower_thresh = 3,
+                     max.em.its = 1000)
+
+saveRDS(CAA18_stm, here::here("analysis/data/derived-data/CAA18_topic_model_and_plot.rds"))
+
+EAA18_stm <- 
+  stm_model_and_plot(conf_paper_titles = EAA18, 
+                     conf_gender_results = EAA18_gm, 
+                     first_name_appears_first = FALSE,
+                     max.em.its = 1000)
+
+saveRDS(EAA18_stm, here::here("analysis/data/derived-data/EAA18_topic_model.rds"))
+
 ```
-
-Figure \@ref(fig:demo-plot) shows how we can have a caption and cross-reference for a plot
-
-```{r demo-inline-code}
-x <- round(pi, 2)
-```
-
-Here is an example of inline code `r x` in the middle of a sentence.
 

### 2020-05-03 analysis/paper/function-for-structural-topic-model.R

@@ -1,0 +1,243 @@
+
+
+stm_plot <- function(conf_paper_titles,
+                     conf_gender_results,
+                     max.em.its = 500,
+                     k = 0,
+                     n_words = 12,
+                     first_name_appears_first = TRUE,
+                     lower_thresh = 15){
+
+  library(here)
+  library(tidyverse)
+  library(readxl)
+  library(quanteda)
+  library(stm)
+  library(dplyr)
+
+  x <- conf_paper_titles
+  d <- conf_gender_results
+
+  x$session <- as.character(x$session)
+  d$session <- as.character(d$session)
+
+
+if (first_name_appears_first) {
+
+  # do nothing to the names
+
+} else {
+
+  # for CAA which has lastname, firstname (affiliation)
+  # extract the first names
+  x$FirstName <- map_chr(x$FirstName,
+                                     ~tolower(str_split(str_split(.x, ".,")[[1]][2],
+                                                        boundary("word"))[[1]][1]))
+}
+
+the_data <- left_join(d, x, by = c("name" = "FirstName",
+                               "session" = "session"))
+
+processed <-
+  textProcessor(the_data$AbstractTitle,
+                metadata = the_data)
+
+out <-
+  prepDocuments(processed$documents,
+                processed$vocab,
+                processed$meta)
+
+docs <- out$documents
+vocab <- out$vocab
+meta <- out$meta
+
+out <-
+  prepDocuments(processed$documents,
+                processed$vocab,
+                processed$meta,
+                lower.thresh = lower_thresh)
+
+set.seed(33)
+
+the_model <<-
+  stm(
+    max.em.its = max.em.its,
+    seed = 33,
+    documents = out$documents,
+    vocab = out$vocab,
+    K = k,
+    # watch the output to see how many topics it picks
+    prevalence =  ~ gender,
+    data = out$meta,
+    init.type = "Spectral"
+  )
+
+# saveRDS(the_model, here::here("analysis/data/derived-data/the_model.rds"))
+# the_model <- readr::read_rds(here::here("analysis/data/derived-data/the_model.rds"))
+
+# if(exists("the_model")){
+#   message("the model object exists")
+# } else {
+#   message("where is the model object?")
+# }
+
+# print(str(the_model))
+
+topic_labels <-
[…]

### 2020-05-03 analysis/paper/functions-for-gender.R

@@ -2,7 +2,7 @@
 
 # read in data
 
-get_the_gender_data <- function(x){
+get_the_gender_data <- function(x, first_name_appears_first = TRUE){
 
   d <- x
 
@@ -71,10 +71,21 @@ require("Hmisc")
 library(purrr)
 first_word_safe <- safely(first.word)
 
+if(first_name_appears_first){
+
 d$first_name <-
-  map(d$FirstName, first_word_safe) %>%
+  map(d$`first-named speaker`, first_word_safe) %>%
   map(., c(1)) %>%
   flatten_chr()
+
+} else {
+
+  # for CAA which has lastname, firstname (affiliation)
+  d$`first-named speaker` <- map_chr(d$`first-named speaker`,
+    ~tolower(str_split(str_split(.x, ".,")[[1]][2],
+                      boundary("word"))[[1]][1]))
+
+}
 
 # fill down on session numbers because I only numbered the first speaker in each session.
 library(tidyr)
@@ -84,7 +95,7 @@ d <- fill(d, session)
 d <-
   d %>%
   group_by(session) %>%
-  distinct(FirstName, .keep_all = TRUE )
+  distinct(`first-named speaker`, .keep_all = TRUE )
 
 # compute gender of first name
 library(gender)

### 2020-05-03 analysis/paper/structural-topic-model.R

@@ -1,182 +1,0 @@
-library(here)
-library(tidyverse)
-library(readxl)
-library(quanteda)
-x <-
-  readxl::read_excel(here("analysis/data/raw-data/2018 SAA Title.xlsx"))
-
-library(stm)
-library(dplyr)
-
-d <- readRDS(here::here("analysis/data/derived-data/gender unnest"))
-
-data <- left_join(d, x, by = c("name" = "FirstName",
-                               "SessionNumber" = "session"))
-
-processed <-
-  textProcessor(data$AbstractTitle,
-                metadata = data)
-
-out <-
-  prepDocuments(processed$documents,
-                processed$vocab,
-                processed$meta)
-
-docs <- out$documents
-vocab <- out$vocab
-meta <- out$meta
-
-out <-
-  prepDocuments(processed$documents,
-                processed$vocab,
-                processed$meta,
-                lower.thresh = 15)
-
-set.seed(33)
-model <-
-  stm(
-    seed = 33,
-    documents = out$documents,
-    vocab = out$vocab,
-    K = 0,
-    # watch the output to see how many topics it picks
-    prevalence =  ~ gender,
-    data = out$meta,
-    init.type = "Spectral"
-  )
-
-topic_labels <-
-  data.frame(labelTopics(model, n = 12)$frex)
-
-topic_labels_chr <-
-topic_labels %>%
-  t %>%
-  as_tibble() %>%
-  map(function(...) {
-    current <- tibble(...)
-     current  %>%
-       mutate_all(as.character) %>%
-       unlist(., use.names=FALSE) %>%
-       paste0(., collapse = ", ")
-  }) %>%
-  unlist %>%
-  unname
-
-out$meta$gender <- as.factor(out$meta$gender)
-
-prep <- estimateEffect(1:model$settings$dim$K ~ gender,
-                       model,
-                       meta = out$meta,
-                       uncertainty = "Global")
-
-topic <-
-  data.frame(
-    topicnames = topic_labels_chr,
-    TopicNumber = 1:model$settings$dim$K,
-    TopicProportions = colMeans(model$theta),
-    stringsAsFactors = F
-  )
-
-Result <-
-plot(
-  prep,
-  method = "difference",
-  covariate = "gender",
-  cov.value1 = "female",
-  cov.value2 = "male",
-  model = model,
-  omit.plot = TRUE
-)
-
-
-trank = order(unlist(Result$means))
-temp.topic <- topic[trank, ]
-
-Result <-
-  plot(
-    prep,
-    method = "difference",
[…]

### 2020-05-03 analysis/paper/paper.Rmd
we have plots for all three confs

@@
 
-saveRDS(EAA18_stm, here::here("analysis/data/derived-data/EAA18_topic_model.rds"))
+saveRDS(EAA18_stm, here::here("analysis/data/derived-data/EAA18_topic_model_and_plot.rds"))
+```
 
+```{r}
+# get plots
+
+# SAA -----
+SAA18_stm_plot <- 
+  readr::read_rds(here::here("analysis/data/derived-data/SAA18_topic_model_and_plot.rds"))
+
+SAA18_stm_plot$plot
+ggsave(here::here("analysis/figures/SAA18_stm_plot.png"), height = 7, width = 6)
+
+# CAA -----
+CAA18_stm_plot <- 
+  readr::read_rds(here::here("analysis/data/derived-data/CAA18_topic_model_and_plot.rds"))
+
+CAA18_stm_plot$plot
+ggsave(here::here("analysis/figures/CAA18_stm_plot.png"), height = 7, width = 6)
+
+# EAA -----
+EAA18_stm_plot <- 
+  readr::read_rds(here::here("analysis/data/derived-data/EAA18_topic_model_and_plot.rds"))
+
+EAA18_stm_plot$plot
+ggsave(here::here("analysis/figures/EAA18_stm_plot.png"), height = 7, width = 6)
 ```

### 2020-05-04 analysis/paper/paper.Rmd
Results try 1

@@
 
-- some basic details ( 1 para)
--- number of papers that are in our sample for each conference
--- number of men and women first-named presented, state the gender ratio 
--- number of people that we could not assign a gender to
+In our analysis we have a total paper distribution of SAA with `SAA18_row` papers, EAA with `EAA18_row` papers and CAA with `CAA18_row` of papers.
+For SAA, we have `female_count_SAA18` females and `male_count_SAA18` males, for EAA, we have `female_count_EAA18` females and `male_count_EAA18` males, finally for CAA, we have `female_count_CAA18` females and `male_count_CAA18` males.
+This means that our ratio of females to males in each of the cnferences are as follows, SAA with `female_to_male_SAA18`, EAA with `female_to_male_EAA18` and CAA with `female_to_male_CAA18`.
+As noted earlier out data set did not originally include the gender of each presenters' thus we are only able to assign gender to presenters based on their first names, and this means that some people are not able to receive a gender due to our filter, in SAA we are forced to drop `missing_gender_SAA18` number of topics,
+`missing_gender_EAA18` topics in EAA, `missing_gender_CAA18` topics in CAA.
 
-The number of ... `r number_of_papers_saa_2018` ...
+We have used the built in function of the R package stm to calculate the number of topics that is best for each conferences, for SAA we have generated a total of `SAA_stm_topic_number` topics, for EAA we have generated a total of `EAA_stm_topic_number` topics, and for CAA we hve generated a total of `CAA_stm_topic_number`.
 
-- the main details
--- number of topics that are optimum for each conf (1 para)
--- variance of topics with gender for each conf (1 para)
+First let's examine the figure for SAA 2018's stm topic plot that tells us where a topic seems to fall under.
+One of the main difference that we can see between the topics of the genders are the locations that the presenters are interested in, it seems like that females are more likely to have a topic at a specific region such as mesoamerica, arizona and national parks, while males are more likely to have a topic that are in the western regions.
+Next, let's take a look at EAA's difference between females and males.
+Although both red-marked topics include topics with "mediterranean", it is important to distinguish that females seems to be focusing on western mediterranean while males seems to be focusing on northern mediterranean.
+Another main difference that we have noticed is that males tends to have topics that are related to architect, while the topic regarding research methods using isotopes seems to be more popular for females.
+Lastly, in CAA's stm topic plot we can see a clearer pattern from the topics marked in red.
+For females from CAA, we see that topics that are associated with Geographic Information System (GIS) are more likely to be a topic for female presenters.
+Another female-associated topics from the CAA topic plot is topics that includes the word women, excav and inform.
+As for males from CAA, there doesn't seem to have a specific word that is more associated with males besides the word enviorn.
 
@@
 saveRDS(CAA18_gm, here::here("analysis/data/derived-data/CAA18_gender_model.rds"))
+```
+
+```{r, eval=FALSE}
+SAA18_gender <-
+  readRDS(here::here("analysis/data/derived-data/SAA18_gender_model.rds"))
+EAA18_gender <-
+  readRDS(here::here("analysis/data/derived-data/EAA18_gender_model.rds"))
+CAA18_gender <-
+  readRDS(here::here("analysis/data/derived-data/CAA18_gender_model.rds"))
+
+female_count_SAA18 <- 
+  length(which(SAA18_gender$gender == "female"))
+male_count_SAA18 <- 
+  length(which(SAA18_gender$gender == "male"))
+female_to_male_SAA18 <- 
+  female_count_SAA18 / male_count_SAA18
+missing_gender_SAA18 <- 
+  SAA18_row - (female_count_SAA18 + male_count_SAA18)
+
+female_count_EAA18 <- 
+  length(which(EAA18_gender$gender == "female"))
+male_count_EAA18 <- 
+  length(which(EAA18_gender$gender == "male"))
+female_to_male_EAA18 <- 
+  female_count_EAA18 / male_count_EAA18
+missing_gender_EAA18 <-
+  EAA18_row - (female_count_EAA18 + male_count_EAA18)
+
+female_count_CAA18 <- 
+  length(which(CAA18_gender$gender == "female"))
+male_count_CAA18 <- 
+  length(which(CAA18_gender$gender == "male"))
+female_to_male_CAA18 <- 
+  female_count_CAA18 / male_count_CAA18
+missing_gender_CAA18 <-
+  CAA18_row - (female_count_CAA18 + male_count_CAA18)
 ```
@@
 
+SAA_stm_topic_number <-
+  SAA18_stm_plot$model$settings$dim$K
+
 # CAA -----
@@
 
+CAA_stm_topic_number <-
+  CAA18_stm_plot$model$settings$dim$K
+
 # EAA -----
@@
 ggsave(here::here("analysis/figures/EAA18_stm_plot.png"), height = 7, width = 6)
+
+EAA_stm_topic_number <-
+  EAA18_stm_plot$model$settings$dim$K
 ```

### 2020-05-05 analysis/paper/paper.Rmd
read in gender data
editing text

--- 2020-05-03 analysis/paper/paper.Rmd
@@
 output: 
-    bookdown::word_document2:
+    bookdown::html_document2:
       fig_caption: yes
@@
 # Results
-```{r}
-number_of_papers_saa_2018 <-  ...
-```
-
-- some basic details ( 1 para)
--- number of papers that are in our sample for each conference
--- number of men and women first-named presented, state the gender ratio 
--- number of people that we could not assign a gender to
-
-The number of ... `r number_of_papers_saa_2018` ...
-
-- the main details
--- number of topics that are optimum for each conf (1 para)
--- variance of topics with gender for each conf (1 para)
 
@@
 source(here::here("analysis/paper/function-for-structural-topic-model.R"))
+
+SAA18_gm <- readr::read_rds(here::here("analysis/data/derived-data/SAA18_gender_model.rds")) 
+EAA18_gm <- readr::read_rds(here::here("analysis/data/derived-data/EAA18_gender_model.rds"))
+CAA18_gm <- readr::read_rds(here::here("analysis/data/derived-data/CAA18_gender_model.rds"))
 
@@
 # get plots
+library(tidyverse)
 
@@
 SAA18_stm_plot$plot
-ggsave(here::here("analysis/figures/SAA18_stm_plot.png"), height = 7, width = 6)
+ggplot2::ggsave(here::here("analysis/figures/SAA18_stm_plot.png"), 
+       height = 7, 
+       width = 6)
+
+SAA_stm_topic_number <-
+  SAA18_stm_plot$model$settings$dim$K
+SAA_stm_topic_number_significant_covary <- 
+  nrow(layer_data(SAA18_stm_plot$plot, 3))
 
@@
 
+CAA_stm_topic_number <-
+  CAA18_stm_plot$model$settings$dim$K
+CAA_stm_topic_number_significant_covary <- 
+  nrow(layer_data(CAA18_stm_plot$plot, 3))
+
 # EAA -----
@@
 ggsave(here::here("analysis/figures/EAA18_stm_plot.png"), height = 7, width = 6)
+
+EAA_stm_topic_number <-
+  EAA18_stm_plot$model$settings$dim$K
+EAA_stm_topic_number_significant_covary <- 
+  nrow(layer_data(EAA18_stm_plot$plot, 3))
 ```
+
+```{r, eval=TRUE}
+SAA18_gender <-
+  readRDS(here::here("analysis/data/derived-data/SAA18_gender_model.rds"))
+EAA18_gender <-
+  readRDS(here::here("analysis/data/derived-data/EAA18_gender_model.rds"))
+CAA18_gender <-
+  readRDS(here::here("analysis/data/derived-data/CAA18_gender_model.rds"))
+
+count_gendered_SAA18 <- nrow(SAA18_gender)
+perc_gendered_SAA18 <- round(nrow(SAA18_gender) / SAA18_row, 2) * 100
+female_count_SAA18 <- 
+  length(which(SAA18_gender$gender == "female"))
+male_count_SAA18 <- 
+  length(which(SAA18_gender$gender == "male"))
+female_to_male_SAA18 <- 
+  round(female_count_SAA18 / male_count_SAA18,1)
+missing_gender_SAA18 <- 
+  SAA18_row - (female_count_SAA18 + male_count_SAA18)
+
+count_gendered_EAA18 <- nrow(EAA18_gender)
+perc_gendered_EAA18 <- round(nrow(EAA18_gender) / EAA18_row, 2) * 100
+female_count_EAA18 <- 
+  length(which(EAA18_gender$gender == "female"))
+male_count_EAA18 <- 
+  length(which(EAA18_gender$gender == "male"))
+female_to_male_EAA18 <- 
+  round(female_count_EAA18 / male_count_EAA18, 1)
+missing_gender_EAA18 <-
+  EAA18_row - (female_count_EAA18 + male_count_EAA18)
+
+count_gendered_CAA18 <- nrow(CAA18_gender)
+perc_gendered_CAA18 <- round(nrow(CAA18_gender) / CAA18_row, 2) * 100
+female_count_CAA18 <- 
+  length(which(CAA18_gender$gender == "female"))
+male_count_CAA18 <- 
+  length(which(CAA18_gender$gender == "male"))
+female_to_male_CAA18 <- 
+  round(female_count_CAA18 / male_count_CAA18, 1)
+missing_gender_CAA18 <-
+  CAA18_row - (female_count_CAA18 + male_count_CAA18)
+```
+
+In the SAA data we found at total of `r SAA18_row` presentations.
+Of these we could identify the first-named author of `r count_gendered_SAA18` (`r perc_gendered_SAA18`%) presentations as either men (n=`r male_count_SAA18`) or women (n=`r female_count_SAA18`).
+For the CAA meeting we have `r CAA18_row` presentations, and `r count_gendered_CAA18` (`r perc_gendered_CAA18`%) first-named authors could be classified as either men (n=`r male_count_CAA18`) or women (n=`r female_count_CAA18`).
+For the EAA data we have `r EAA18_row` presentations, and `r count_gendered_EAA18` (`r perc_gendered_EAA18`%) where the first author could be classified as either a man (n=`r male_count_EAA18`) or woman (n=`r female_count_EAA18`).
+This results in similar women-to-men ratios for the SAA (`r female_to_male_SAA18`) and EAA (`r female_to_male_EAA18`), but much lower for the CAA (`r female_to_male_CAA18`).
+
+In the SAA presentations we identified the optimum number of topics as `r SAA_stm_topic_number`, with `r SAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
+One of the main difference that we can see between the topics of the genders are the locations that the presenters are interested in, it seems like that females are more likely to have a topic at a specific region such as mesoamerica, arizona and national parks, while males are more likely to have a topic that are in the western regions.
+
+For EAA we have generated a total of `r EAA_stm_topic_number` topics, with `r EAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
+Next, let's take a look at EAA's difference between females and males.
+Although both red-marked topics include topics with "mediterranean", it is important to distinguish that females seems to be focusing on western mediterranean while males seems to be focusing on northern mediterranean.
+Another main difference that we have noticed is that males tends to have topics that are related to architect, while the topic regarding research methods using isotopes seems to be more popular for females.
+
+For CAA we hve generated a total of `r CAA_stm_topic_number`, with `r CAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter..
+Lastly, in CAA's stm topic plot we can see a clearer pattern from the topics marked in red.
+For females from CAA, we see that topics that are associated with Geographic Information System (GIS) are more likely to be a topic for female presenters.
+Another female-associated topics from the CAA topic plot is topics that includes the word women, excav and inform.
+As for males from CAA, there doesn't seem to have a specific word that is more associated with males besides the word enviorn.
 

### 2020-05-08 analysis/paper/paper.Rmd
changed a word
citation

@@
 
-Here is a citation [@Marwick2017]
-
-Gender under-representation and inequality has long existed, from under-representation in children’s picture books to academic publications (Hamilton et al.
-2006; Tushingham et al.
-2017).
+Gender under-representation and inequality has long existed, from under-representation in children’s picture books to academic publications (@hamilton2006gender; @tushingham2017peer).
 When talking about inequality in academic fields, we usually refer to the disproportionality between men and women in the field, including archaeology.
-Women are often underrepresented due to the lack of support women receive in academic fields (Xu 2008).
-When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field (Bardolph et al.
-2016).
-As we can see in the authorship of the main authors from the Journal of Field Archaeology, 78% of them are men (Heath-Stout 2020).
+Women are often underrepresented due to the lack of support women receive in academic fields (@xu2008gender).
+When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field (@bardolph2014critical).
+As we can see in the authorship of the main authors from the Journal of Field Archaeology, 78% of them are men (@heath2020guest).
 When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, Tushingham et al.
@@
 We obtained publicly available programs of these conferences and also tidied the data into a rectangular form with variables including the primary author's first name, last name and the title of their presentation.
-We used R and the gender package (Mullen 2018; Mihaljevic 2019; Belvis et al.
-2015) to predict the gender of all first-named presenters in our sample.
-We used the gender package to identify a first name as male or female, based on a probability scale computed from government records of name-gender data, specifically US Social Security Administration (SSA) baby name data (Mullen 2019).
+We used R and the gender package (@lincolngender; @mihaljevic2019reflections; @blevins2015jane) to predict the gender of all first-named presenters in our sample.
+We used the gender package to identify a first name as male or female, based on a probability scale computed from government records of name-gender data, specifically US Social Security Administration (SSA) baby name data (@mihaljevic2019reflections).
 This method has the advantage of speed and ease of automation, but also some substantial limitations.
@@
 This process "attaches" the metadata to the collection of texts that we will then perform topic modeling process.
-It is important to note that we have used the stm package that also includes the ability to calculate the number of topics that is best suited for STM modeling (Roberts 2019).
-The fact that we are allowed to do this allows us to perform covariate analysis (Roberts et al.
-2013).
+It is important to note that we have used the stm package that also includes the ability to calculate the number of topics that is best suited for STM modeling (@roberts2019package).
+The fact that we are allowed to do this allows us to perform covariate analysis.
 This basically means that we can examine the relationship between the topics and some covariates/variables of interest.
-We can define the genders of males and females and then analyze the prevelance of these two categories in accordance with the topics that the stm package determined (Roberts 2019).
+We can define the genders of males and females and then analyze the prevelance of these two categories in accordance with the topics that the stm package determined (@roberts2019package).
 The analysis is used also with probabilities given to the texts to see if the most occuring of some text seems to be "attached" to one gender over another.
@@
 
-In the SAA data we found at total of `r SAA18_row` presentations.
+In the SAA data we found a total of `r SAA18_row` presentations.
 Of these we could identify the first-named author of `r count_gendered_SAA18` (`r perc_gendered_SAA18`%) presentations as either men (n=`r male_count_SAA18`) or women (n=`r female_count_SAA18`).

### 2020-05-09 analysis/paper/paper.Rmd
editing text

@@
 
-Gender under-representation and inequality has long existed, from under-representation in children’s picture books to academic publications (@hamilton2006gender; @tushingham2017peer).
-When talking about inequality in academic fields, we usually refer to the disproportionality between men and women in the field, including archaeology.
-Women are often underrepresented due to the lack of support women receive in academic fields (@xu2008gender).
-When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field (@bardolph2014critical).
-As we can see in the authorship of the main authors from the Journal of Field Archaeology, 78% of them are men (@heath2020guest).
-When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, Tushingham et al.
-(2017) report a trend of women increasingly maintaining their society membership and the gender gap is almost gone.
-In this study we investigate gender ratios of presenters at major archaeological conferences.
-We extend the work of Tushingham et al.
-(2017) with a study of gender ratios of presenters at meetings of the Society of American Archaeology (SAA, 2016-2019), the European Association of Archaeologists (EAA, 2018) and the Computer Applications and Quantitative Methods in Archaeology (CAA, 2017)  We explore how gender ratios vary over time and between these conferences.
-We examine the influence of the gender of the presenter on the topics that they present on.
+Gender under-representation and inequality has long existed in science, from under-representation of women and minority genders in school text books to academic publications (@hamilton2006gender; @tushingham2017peer).
+Women are often under-represented for several reasons, including stereotyping that unfairly underestimates individuals' abilities, which often leads to lack of support for women in academic fields (@xu2008gender).
+In this paper we explore the relationship between gender and research topics to determine if women and men tend to focus more on certain topics than others.
+
+We used a computational method to identify topics from the titles of presentations delivered at meetings of the Society of American Archaeology (SAA, 2016-2019), the European Association of Archaeologists (EAA, 2018) and the Computer Applications and Quantitative Methods in Archaeology (CAA, 2017).
+We analysed the covariance of topics and genders of presenters to explore how gender ratios vary between these conferences and by topics.
 
 # Background
+
+When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field (@bardolph2014critical).
+
+As we can see in the authorship of the main authors from the Journal of Field Archaeology, 78% of them are men (@heath2020guest).
+
+When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership and the gender gap is almost gone.
+In this study we investigate gender ratios of presenters at major archaeological conferences.
 
@@
 
+library(gender)
+lynne_gender <-  gender("Lynne")
+
 ```
@@
 
-We collected lists of the people that presented research at recent SAA (2018 with `r SAA18_row` presenters, EAA (2018 with `r EAA18_row` presenters) and CAA (2018 with `r CAA18_row` presenters) conferences.
-We obtained publicly available programs of these conferences and also tidied the data into a rectangular form with variables including the primary author's first name, last name and the title of their presentation.
-We used R and the gender package (@lincolngender; @mihaljevic2019reflections; @blevins2015jane) to predict the gender of all first-named presenters in our sample.
-We used the gender package to identify a first name as male or female, based on a probability scale computed from government records of name-gender data, specifically US Social Security Administration (SSA) baby name data (@mihaljevic2019reflections).
-This method has the advantage of speed and ease of automation, but also some substantial limitations.
-We are only able to infer gender into binary male/female genders and assign probabilities for the first names into these categories.
-This has the unfortunate result of excluding other genders from the results, and we encourage conference organizers to collect gender data directly from participants to improve the representation of minority genders in future studies.
-The package often fails to classify non-English names, as the SSA data the package uses are mostly in English names.
+We requested from conference organisers spreadsheets files of the publicly available program information for three major archaeology meetings held in 2018: the Society of American Archaeology (SAA, 2016-2019), the European Association of Archaeologists (EAA, 2018) and the Computer Applications and Quantitative Methods in Archaeology (CAA, 2017).
+
+We estimated the gender of the first-named speaker for each presentation using the R programming language and the gender package (@lincolngender; @mihaljevic2019reflections; @blevins2015jane).
+We inferred the gender of personal names by looking each name up in the US Social Security Administration (SSA) baby name data, and calculating the overall probability that a given name was male or female (@mihaljevic2019reflections).
+If the proportion of people in the SSA data with a certain given name are recorded as female is 0.5 or higher, we predicted the gender of the presenter with this name as female.
+For example, in the SSA dataset the the name "Lynne" returns `r lynne_gender$proportion_male` as proportion of individuals with this name who are recorded as male, and `r lynne_gender$proportion_female` as the proportion female, thus we inferred that people in our conference data who are called Lynne are women.
+
+This method of inferring gender has the advantage of speed, transparency and reproducibility, but also some substantial limitations that are important to be upfront about.
+We are only able to infer binary male/female genders and assign the first names into these two categories.
+This has the unfortunate result of excluding other genders from the results, excluding them from our analysis.
+We considered it impractical and invasive to write to each first-named presented to request their gender information.
+A further limitation of our approach is that it sometimes fails to classify non-English names at all, as the SSA data consists mostly of English names.
 This means that people with non-English names are underrepresented in our results.
-To remove names where gender is ambiguous, which we define as a similar proportion of males and females having that name in the SSA data, we excluded names where the difference in proportions was less than 0.47, a figure we obtained by exploratory data analysis.
-It is important to note that the gender designations here are not self-identified by the presenters, but are computed probabilistically.
-Better quality data will come if presenters self-identify their gender, but we have no alternative because currently these data are not available.
+It is important to note that the inferences of gender presented here are not self-identified by the presenters, but are computed probabilistically.
+Better quality and more representative data would result if presenters self-identified their gender, but currently these data are not available.
+We encourage conference organizers to collect gender data directly from participants to improve the representation of minority genders in future studies.
 Our hope is that this work, despite its many limitations, will stimulate the collection of more reliable, justifiable, and useful data by conference organisers.
 
-We are also interested in whether there is a variation in topic choices between these two genders in archaeological conferences.
-To do this we have used topic modeling, which is a method to automatically find related groups of words that resemble 'themes' or 'topics'.
-This approach does not treat one topic, which consists of many words, as separate from another topic.
-That is, a word can be present in many topics but with differs weightings.
+We identified the topic of each presentation by generating a structural topic model for all the presentations in each of the three conferences.
+Topic modeling is a machine learning method to automatically find related groups of words that resemble traditional 'themes' or 'topics' in a collection of documents.Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields [add citations].
+This approach uses  latent Dirichlet allocation to allow every word to be present in every topic, but with different weightings, such that the most heavily weighted 5-10 words of a topic often capture the essence of the topic as a coherent and familiar concept.
 When generating topic models, we have to first decide on the number of topics that the method will identify in our texts.
-We then create a topic model for all our texts, this will assign all words in every document into our set number of topics, and assign each word a probability based on its per-topic-per-word probability.
-To visualise the topic model we use the top 10 words within each topic as keywords to represent each topic.
-We excluded words that are common in the field of archaeology or have little semantic value, for example, 'the', 'a', 'el', 'al', 'archaeology', etc.
-We generated one topic model per conference.
-Each presentation at each conference is then assigned a vector of topic probabilities.
+The topic modeling algorithm then  assigns all words in every document into our set number of topics, and assigns each word a probability based on its per-topic-per-word probability, and each document (i.e.
+conference presentation title) is assigned a distribution of topics with different weights.
+To visualise the topic model output we use the 10 words with the highest weighting within each topic as keywords to represent each topic.
+In generating the topic model we excluded words that are common in English generally or in archaeology and so have little semantic value, for example, 'the', 'a', 'el', 'al', 'archaeology', etc.
 
-We then tried another process of topic modeling called Structural Topic Model (STM) which is very similar to the LDA we have described above.
-The main advantage for us is that, the STM allows us to include, alongside the normal topic modeling process, other metadata.
-This process "attaches" the metadata to the collection of texts that we will then perform topic modeling process.
-It is important to note that we have used the stm package that also includes the ability to calculate the number of topics that is best suited for STM modeling (@roberts2019package).
-The fact that we are allowed to do this allows us to perform covariate analysis.
-This basically means that we can examine the relationship between the topics and some covariates/variables of interest.
-We can define the genders of males and females and then analyze the prevelance of these two categories in accordance with the topics that the stm package determined (@roberts2019package).
-The analysis is used also with probabilities given to the texts to see if the most occuring of some text seems to be "attached" to one gender over another.
-
-To explore the relationship between gender and topics we identified the most prominent topic for each presentation...
+A structural topic model (STM) allows us to include, alongside the typical topic modeling process, other document-level metadata to analyse covariance between topics and metadata (@roberts2019package).
+Covariates of interest can be included into the prior distributions for document-topic proportions and
+topic-word distributions using a standard regression model (cite this one: http://scholar.harvard.edu/files/dtingley/files/topicmodelsopenendedexperiments.pdf).
+This means that we can examine the relationship between the topics and document-level variables of interest.
+In our case, we computed the relationship between the inferred the gender of the first-named presenter for each presentation, and the distribution of topics in each presentation.
+When we observed the topics for each presentation, we generated a regression where the topic is the outcome variable, and gender is the explanatory variable.
+This regression gives us insights into whether gender caused archaeologists to spend a larger portion of their presentation on a particular topic.
 
@@
 
-In the SAA data we found a total of `r SAA18_row` presentations.
+In the SAA data we found a total of `r prettyNum(SAA18_row, big.mark = ",")` presentations.
 Of these we could identify the first-named author of `r count_gendered_SAA18` (`r perc_gendered_SAA18`%) presentations as either men (n=`r male_count_SAA18`) or women (n=`r female_count_SAA18`).
-For the CAA meeting we have `r CAA18_row` presentations, and `r count_gendered_CAA18` (`r perc_gendered_CAA18`%) first-named authors could be classified as either men (n=`r male_count_CAA18`) or women (n=`r female_count_CAA18`).
-For the EAA data we have `r EAA18_row` presentations, and `r count_gendered_EAA18` (`r perc_gendered_EAA18`%) where the first author could be classified as either a man (n=`r male_count_EAA18`) or woman (n=`r female_count_EAA18`).
+For the CAA meeting we have `r prettyNum(CAA18_row, big.mark = ",")` presentations, and `r count_gendered_CAA18` (`r perc_gendered_CAA18`%) first-named authors could be classified as either men (n=`r male_count_CAA18`) or women (n=`r female_count_CAA18`).
+For the EAA data we have `r prettyNum(EAA18_row, big.mark = ","` presentations, and `r count_gendered_EAA18` (`r perc_gendered_EAA18`%) where the first author could be classified as either a man (n=`r male_count_EAA18`) or woman (n=`r female_count_EAA18`).
 This results in similar women-to-men ratios for the SAA (`r female_to_male_SAA18`) and EAA (`r female_to_male_EAA18`), but much lower for the CAA (`r female_to_male_CAA18`).
@@
 In the SAA presentations we identified the optimum number of topics as `r SAA_stm_topic_number`, with `r SAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
-One of the main difference that we can see between the topics of the genders are the locations that the presenters are interested in, it seems like that females are more likely to have a topic at a specific region such as mesoamerica, arizona and national parks, while males are more likely to have a topic that are in the western regions.
+One of the main difference that we can see between the topics of the genders are the locations that the presenters are interested in.
+Women are more likely to have a topic at a specific region such as Mesoamerica, Arizona and national parks, while men are more likely to have a topic that are in the western regions.
 
-For EAA we have generated a total of `r EAA_stm_topic_number` topics, with `r EAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
-Next, let's take a look at EAA's difference between females and males.
-Although both red-marked topics include topics with "mediterranean", it is important to distinguish that females seems to be focusing on western mediterranean while males seems to be focusing on northern mediterranean.
-Another main difference that we have noticed is that males tends to have topics that are related to architect, while the topic regarding research methods using isotopes seems to be more popular for females.
+For EAA we generated a total of `r EAA_stm_topic_number` topics, with `r EAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
+Although topics  significantly associated with men and women both include the word "Mediterranean", it is important to distinguish that females seems to be focusing on western Mediterranean while males seems to be focusing on northern Mediterranean Another  difference  is that males tends to have topics that are related to architecture, while the topic regarding research methods using isotopes seems to be more popular for females.
 
-For CAA we hve generated a total of `r CAA_stm_topic_number`, with `r CAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter..
-Lastly, in CAA's stm topic plot we can see a clearer pattern from the topics marked in red.
+For CAA we generated a total of `r CAA_stm_topic_number` topics, with `r CAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
+In CAA's stm topic plot we can see a clearer pattern from the topics marked in red.
 For females from CAA, we see that topics that are associated with Geographic Information System (GIS) are more likely to be a topic for female presenters.

### 2020-05-09 analysis/paper/functions-for-joining-table.R

@@ -1,62 +1,0 @@
-
-
-# functions for joining table
-
- x <-readRDS(here::here("analysis/data/derived-data/SAA18_topic_model.rds"))
- y <-readRDS(here::here("analysis/data/derived-data/SAA18_gender_model.rds"))
-
-joining_table_topic_and_gender <- function(x, y)
-
-library(dplyr)
-library(tidyverse)
-joined_table <-
-  left_join(x$unique_id_gamma,
-            y,
-            by = c("FirstName" = "name",
-                   "session" = "session"))
-
-
-#Making table similar to Stats Rethinking
-
-
-
-joined_table %>%
-  filter(!is.na(gender)) %>%
-  filter(gamma >= 0.1) %>%
-ggplot(aes(x = as.factor(topic), y = gamma)) +
-  geom_boxplot() +
-  facet_wrap(~gender, ncol = 1)
-
-
-joined_table %>%
-  filter(!is.na(gender)) %>%
-  filter(gamma >= 0.1) %>%
-  mutate(new_gamma = ((gamma - mean(gamma)) / sd(gamma))^(1/2)) %>%
-  ggplot(aes(x = as.factor(topic),
-             y = new_gamma,
-             colour = gender)) +
-  geom_boxplot()
-
-
-joined_table %>%
-  filter(!is.na(gender)) %>%
-  filter(gamma >= 0.1) %>%
-  mutate(new_gamma = ((gamma - mean(gamma)) / sd(gamma))^(1/2)) %>%
-  filter(!is.nan(new_gamma)) %>%
-  mutate_at(vars(topic, gender),
-            as.factor) %>%
-  aov(gamma ~ gender * topic, data = . ) %>%
-  summary
-
-d <-
-  joined_table %>%
-  filter(!is.na(gender)) %>%
- # filter(gamma >= 0.1) %>%
-  mutate(new_gamma = ((gamma - mean(gamma)) / sd(gamma))^(1/2)) %>%
-  filter(!is.nan(new_gamma)) %>%
-  mutate(topic = paste0('topic', topic)) %>%
-  select(-gamma) %>%
-  pivot_wider(names_from = topic,
-              values_from = new_gamma,
-              values_fill = list(new_gamma = 0))
-

### 2020-05-09 analysis/paper/functions-for-topic-models.R

@@ -1,78 +1,0 @@
-
-
-# function to compute topic model for a conf
-
-# x is the data frame from the spreadsheet
-
-do_the_topic_model <- function(x, k) {
-
-  d <- x
-
-  library(tidyverse)
-  library(tidytext)
-
-  d_word <- d %>%
-    unnest_tokens(word, AbstractTitle)
-
-  word_counts <- d_word %>%
-    anti_join(stop_words) %>%
-    filter(!word %in% c("de", "la", "el", "los", "user",
-                        "archaeology", "archaeological"
-                        ,"site", "archaic", "sites")) %>%
-    filter(!str_detect(word, "\\d")) %>%
-    filter(str_length(word) >= 3) %>%
-    filter(!str_detect(word, "\\.|'|\\<U")) %>%
-    unite(unique_id, session, FirstName) %>%
-    count(unique_id, word, sort = TRUE) %>%
-    ungroup()
-
-  unique_id_dtm <- word_counts %>%
-    cast_dtm(unique_id, word, n)
-
-  library(topicmodels)
-  unique_id_lda <- LDA(unique_id_dtm,
-                       k = k,
-                       control = list(seed = 1234))
-
-  top_terms <-  tidy(unique_id_lda,
-                     matrix = "beta")  %>%
-    group_by(topic) %>%
-    top_n(10, beta) %>%
-    ungroup() %>%
-    arrange(topic, -beta)
-
-  unique_id_gamma <- tidy(unique_id_lda, matrix = "gamma")
-
-  unique_id_gamma <- unique_id_gamma %>%
-    separate(document, c("session", "FirstName"),
-             sep = "_",
-             convert = TRUE)
-
-  return(list(unique_id_dtm = unique_id_dtm,
-              unique_id_lda = unique_id_lda,
-              top_terms = top_terms,
-              unique_id_gamma = unique_id_gamma))
-
-}
-
-
-
-# access the different elements with $
-
-# compute the optimum number of topics
-
-compute_optimum_n_topics <- function(dtm){
-  ldatuning::FindTopicsNumber(
-    dtm,
-   topics = seq(from = 2, to = 500, by = 1),
-   metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
-   method = "Gibbs",
-   control = list(seed = 77),
-   mc.cores = 2L,
-   verbose = TRUE
- )
-}
-
[…]

### 2020-05-15 analysis/paper/paper.Rmd
some more background

@@
 
-Gender under-representation and inequality has long existed in science, from under-representation of women and minority genders in school text books to academic publications (@hamilton2006gender; @tushingham2017peer).
-Women are often under-represented for several reasons, including stereotyping that unfairly underestimates individuals' abilities, which often leads to lack of support for women in academic fields (@xu2008gender).
+Gender under-representation and inequality has long existed in science, from under-representation of women and minority genders in school text books to academic publications [@hamilton2006gender; @tushingham2017peer].
+Women are often under-represented for several reasons, including stereotyping that unfairly underestimates individuals' abilities, which often leads to lack of support for women in academic fields [@xu2008gender].
 In this paper we explore the relationship between gender and research topics to determine if women and men tend to focus more on certain topics than others.
@@
 
-When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field (@bardolph2014critical).
+When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field [@bardolph2014critical].
+In a previous study done on gender balance of the production of the Atapuerca Project, a project about human evolution in the field of archaeo-palaeontology, they have found that among all the authors, only 35.9% of them are women [@rodriguez2018gender].
+In our study, we also continue the study of gender distribution of presenters at major archaeological conferences.
 
-As we can see in the authorship of the main authors from the Journal of Field Archaeology, 78% of them are men (@heath2020guest).
+Another similar imbalance can be seen in terms of the authorship of articles or papers written.
+For example, for the authorship of the main authors from the Journal of Field Archaeology, 78% of them are men.
+[@heath2020guest].
+This can also be seen in the study of the Atapuerca Project, in which only 34.8% of the papers are written by a female first/lead author [@rodriguez2018gender].
+We have extended the study of ratios of authorships in the field of archaeology by specifically focusing on the gender imbalance that occurs in the topics that the authors present in major archaeological conferences.
 
 When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership and the gender gap is almost gone.
-In this study we investigate gender ratios of presenters at major archaeological conferences.
+Although this similar trend is also present in the study by @vanderwarker2018ucsb in terms of more and more women pursuing a degree in archaeology than men, it is important to note that data collected from University of California, Santa Barbara Gender Equity Project shows that there are still imbalance as a result of different pay rates, professional ranking, and professional advancement [@gonzalez2018perceptions].
 
@@
 
-We estimated the gender of the first-named speaker for each presentation using the R programming language and the gender package (@lincolngender; @mihaljevic2019reflections; @blevins2015jane).
-We inferred the gender of personal names by looking each name up in the US Social Security Administration (SSA) baby name data, and calculating the overall probability that a given name was male or female (@mihaljevic2019reflections).
+We estimated the gender of the first-named speaker for each presentation using the R programming language and the gender package [@lincolngender; @mihaljevic2019reflections; @blevins2015jane].
+We inferred the gender of personal names by looking each name up in the US Social Security Administration (SSA) baby name data, and calculating the overall probability that a given name was male or female [@mihaljevic2019reflections].
 If the proportion of people in the SSA data with a certain given name are recorded as female is 0.5 or higher, we predicted the gender of the presenter with this name as female.
@@
 
-A structural topic model (STM) allows us to include, alongside the typical topic modeling process, other document-level metadata to analyse covariance between topics and metadata (@roberts2019package).
+A structural topic model (STM) allows us to include, alongside the typical topic modeling process, other document-level metadata to analyse covariance between topics and metadata [@roberts2019package].
 Covariates of interest can be included into the prior distributions for document-topic proportions and

### 2020-05-15 analysis/paper/paper.Rmd
working on background section

@@
 When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field [@bardolph2014critical].
-In a previous study done on gender balance of the production of the Atapuerca Project, a project about human evolution in the field of archaeo-palaeontology, they have found that among all the authors, only 35.9% of them are women [@rodriguez2018gender].
-In our study, we also continue the study of gender distribution of presenters at major archaeological conferences.
+@rodriguez2018gender studied gender balance among 309 authors of 299 publications produced during 1978-2016 by the Atapuerca Project, a large archaeo-palaeontological project in Burgos, Spain.
+They 'manually defined' the gender of each author, and found 13 articles authored exclusively by women, 34.8% of papers have women as first-authors, and overall, 35.9% of all authors are women [@rodriguez2018gender].
+Over time, they observed a trend of increasing numbers of papers that include women as authors, numbers of paper lead by women as the first author, and increases in the ratio of female to male authors in group-authored papers.
+They did not make any observations about relationships between gender and topics.
 
 Another similar imbalance can be seen in terms of the authorship of articles or papers written.
-For example, for the authorship of the main authors from the Journal of Field Archaeology, 78% of them are men.
-[@heath2020guest].
-This can also be seen in the study of the Atapuerca Project, in which only 34.8% of the papers are written by a female first/lead author [@rodriguez2018gender].
+For example, for the authorship of the first authors of 1,104 articles published in the _Journal of Field Archaeology_ during 1974–2018, 72% of them are men [@heath2020guest].
 We have extended the study of ratios of authorships in the field of archaeology by specifically focusing on the gender imbalance that occurs in the topics that the authors present in major archaeological conferences.
 
+<!-- change over time -->
 When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership and the gender gap is almost gone.
 Although this similar trend is also present in the study by @vanderwarker2018ucsb in terms of more and more women pursuing a degree in archaeology than men, it is important to note that data collected from University of California, Santa Barbara Gender Equity Project shows that there are still imbalance as a result of different pay rates, professional ranking, and professional advancement [@gonzalez2018perceptions].
+
+<!-- add one more -->
 

### 2020-05-22 analysis/paper/paper.Rmd
l

@@
 <!-- change over time -->
-When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership and the gender gap is almost gone.
-Although this similar trend is also present in the study by @vanderwarker2018ucsb in terms of more and more women pursuing a degree in archaeology than men, it is important to note that data collected from University of California, Santa Barbara Gender Equity Project shows that there are still imbalance as a result of different pay rates, professional ranking, and professional advancement [@gonzalez2018perceptions].
-
+When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership and the gender gap being shortened over the years.
+@tushingham2017peer focuses on 3 articles from journals published in _Journal of California and Great Basin Anthropology/ Journal of California Anrchaeology_ (JCGBA/JCA) and _California Archaeology_ (CA) and a non-peer reviewed Society for California Archaeology _Poroceedings_.
+When looking at the percentage of female lead authors of a total of 2617 authors, the proportion of female lead authors have been increasing statistically significantly with p < 0.01 from the Cochran-Armitage chi-square test for trend (@tushingham2017peer).
+This similar trend of closing the gender gap can also be seen in women's pursue of higher degree in archaeology, out of a total of 217 surveryed, we see that in the 1950s females have 0% that are what the authors considered as decade of highest degree while in the 1980s it grew to 9.7% and eventually to 46.1% in 2010s (@vanderwarker2018ucsb).
+In the same study by @vanderwarker2018uscb, it is important to note that data collected from University of California, Santa Barbara Gender Equity Project shows that there are still imbalance as a result of different pay rates, for example, breaking down annual income by 20000 groupings, there is a trend that women outpoportions men in the lower bracket of less than 20000, 21000 to 40000 and 41000 to 60000 where as men outproportions women in the upper bracket of 61000 to 80000 and 81000 to 100000 and greater than 100000 (@vanderwarker2018uscb).
 <!-- add one more -->
+Something about Twitter
 

### 2020-05-22 analysis/paper/paper.Rmd
l

@@
 
-<!-- change over time -->
 When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership and the gender gap being shortened over the years.
@@
 In the same study by @vanderwarker2018uscb, it is important to note that data collected from University of California, Santa Barbara Gender Equity Project shows that there are still imbalance as a result of different pay rates, for example, breaking down annual income by 20000 groupings, there is a trend that women outpoportions men in the lower bracket of less than 20000, 21000 to 40000 and 41000 to 60000 where as men outproportions women in the upper bracket of 61000 to 80000 and 81000 to 100000 and greater than 100000 (@vanderwarker2018uscb).
-<!-- add one more -->
+
 Something about Twitter

### 2020-05-22 analysis/paper/paper.Rmd
expanding background section

--- 2020-05-15 analysis/paper/paper.Rmd
@@
 When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field [@bardolph2014critical].
+<!--- something about increasing numbers of women's with PhD and increasing rates of participation -->
+
 @rodriguez2018gender studied gender balance among 309 authors of 299 publications produced during 1978-2016 by the Atapuerca Project, a large archaeo-palaeontological project in Burgos, Spain.
@@
 For example, for the authorship of the first authors of 1,104 articles published in the _Journal of Field Archaeology_ during 1974–2018, 72% of them are men [@heath2020guest].
+@heath2020guest investigates the possibility that the gender gap in authorship is due to due to sexism in the peer review process.
+Out of 830 instances of peer review for where both the reviewer’s gender and the first author's gender could be determined, neither the first author’s gender, the reviewer’s gender, nor the combination of the two had a significant effect on the reviewer's recommendation.
+
+@fulkerson2019dominates collected data on author's gender and occupational affiliation in peer reviewed journals (_American Antiquity_ (AQ), _Advances in Archaeological Practice_ (AAP), _Journal of California and Great Basin Anthropology_ (JCGBA), and _California Archaeology_ (CA)) and some non-peer-reviewed venues (the _SAA Archaeological Record_ (SAA Record) and the SCA _Proceedings_).
+Among the 5,010 authors of 2,445 articles in their sample, 27.1% of first/single authors of peer-reviewed journal articles are women, and 72.9% are men.
+The gender gap is less pronounced in the non-peer-reviewed venues with women accounting for 40.8% of lead authors in 517 articles.
+
+When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership they remain underrepresented in peer-reviewed journals.
+@tushingham2017peer examined authorship gender trends in 1,599 papers in three journals, _Journal of California and Great Basin Anthropology/Journal of California Archaeology_ (JCGBA/JCA), _California Archaeology_ (CA) and a non-peer reviewed _Proceedings_ of the Society for California Archaeology (PSCA).
+In a total of 2,617 authors, 844 (32.3%) were women, 1,762 (67.3%) were men, and 11 (0.4%) were gender unknown/ambiguous [@tushingham2017peer].
+They found significant increases in the proportion of female lead authors over time in JCGBA/JCA and PSCA, but not in CA, perhaps because of the shorter publishing period for this journal [@tushingham2017peer].
+This similar trend of a gender gap in archaeology can also be seen in women's pay [@vanderwarker2018ucsb].
+In the University of California, Santa Barbara Gender Equity Project,@vanderwarker2018uscb report an imbalance in pay rates, for example, women make a greater proportion of people paid less \$60,000 and men outnumber women in salary brackets greater than \$60,000.
+(@vanderwarker2018uscb).
+
+<!-- summarise studies of citation practices that show a gap -->
+
+<!-- conference participation, check Bardolph 2018-->
+
+Something about Twitter
+
 We have extended the study of ratios of authorships in the field of archaeology by specifically focusing on the gender imbalance that occurs in the topics that the authors present in major archaeological conferences.
-
-<!-- change over time -->
-When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership and the gender gap is almost gone.
-Although this similar trend is also present in the study by @vanderwarker2018ucsb in terms of more and more women pursuing a degree in archaeology than men, it is important to note that data collected from University of California, Santa Barbara Gender Equity Project shows that there are still imbalance as a result of different pay rates, professional ranking, and professional advancement [@gonzalez2018perceptions].
-
-<!-- add one more -->
 

### 2020-06-05 analysis/paper/paper.Rmd
aaaaa

@@
 
+@bardolph2018controlling has examined data from 2007 to 2017 about the membership in 
+the Society of California Archaeology (SCA) conference, the lead-authors of peer-reviwed journal _Journal of Californial and Great Basin Anthropology_ (JCGBA), and _California Archaeology_ (CA).
+@bardolph2018controlling noted that the women's presentation rate is consistent with their membership rates in SCA, and that women's participation rate in 2017 was finally able to exceed that of men's.
+Although women is more actively participating in SCA, the same is not shown in terms of the lead-authorship of JCGBA and CA [@bardolph2018controlling].
+Throughout the majoirty of time in the period collected, the lead-authorship of JCGBA is obviously skewed towards men and only 34% of them are women.
+A greater difference between men and women lead-authorship can be seen in CA, in 2009 there was no women lead-authors at all and overall only 23% of the publichsed papers of CA are women lead-authors [@bardolph2018controlling].
+
 When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership they remain underrepresented in peer-reviewed journals.
@@
 In the University of California, Santa Barbara Gender Equity Project,@vanderwarker2018uscb report an imbalance in pay rates, for example, women make a greater proportion of people paid less \$60,000 and men outnumber women in salary brackets greater than \$60,000.
-(@vanderwarker2018uscb).
+[@vanderwarker2018uscb].
 
+One common flaw that often exists in the study of genders from lead-authorships is the fact that most are done by the method of predicting the gender based on the author's first name, as noted by @heathwrites, she has recently conducted a survey that directly asks the archaeologists that published in some journals for their self-identifications of gender, race/ethnicity and sexual orientation.
+@heathwrites has used the data she gathered to compare it to previously done studies on similar topic of lead-authorships in jounals such as, _Advances in Archaeological Practice_ (AAP), _American Antiquity_ (AAQ), _Ancient Mesoamerica_ (AM), _Historical Archaeology_ (HA), _Journal of Archaeological Method and Theory_ (JAMT), _Journal of Archaeological Research_ (JAaR), _Journal of Field Archaeology_ (JFA), and _Latin American Antiquity_ (LAQ).
+From the comparison, @heathwrites found that there is a higher number of women publishing in the field of archaeology than it is shown in previous studies in the same publications, which is most likely a result of change overtime, and that most journals is slowly reaching to gender parity.
+
+Inside the field of archaeology there are still a variation of topics to study, which @mallia2009opposite focuses on in her paper using 222 articles from journals of _Revista Espanola de Antropologia Americana_ (REAA) and 206 articles from _Anales del Museo de America_ (AMA), over a time span of 1993 to 2007.
+@mallia2009opposite found that there seems to be some topics that are more commonly studied/published, for example, when dividing the topics into Pre-Columbian, Contact, Colonial and Republican, in REAA there is more collboration with American writers and those that do frequent field work in South America and this corresponded with pre-Columbian as the dominant time span then followed by Republican, while AMA shows a dominant in the time span of Colonial periods which is likely due to the author's study of collections of arts and records from museum.
+Although the article only focused on topics within the field of arhcaeology and its distribution, we are interested in whether gender can be a factor in terms of what kinds of topics are more frequently studied.
 <!-- summarise studies of citation practices that show a gap -->

### 2020-06-05 analysis/paper/paper.Rmd
editing background paras

--- 2020-05-22 analysis/paper/paper.Rmd
@@
 When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field [@bardolph2014critical].
-<!--- something about increasing numbers of women's with PhD and increasing rates of participation -->
+<!--- something about increasing numbers of womens with PhD and increasing rates of participation -->
 
@@
 
+@bardolph2018controlling examined data from 2007 to 2017 about the membership in the Society of California Archaeology (SCA) conference, and the lead-authors of JCGBA and CA.
+@bardolph2018controlling noted that women's conference presentation rates are consistent with their membership rates in SCA, and that it was not until 2017 when women's participation rates finally exceeded men's.
+Although women are more actively participating in SCA, this is not the case for lead-authorship of JCGBA and CA [@bardolph2018controlling].
+Throughout the majority of time in the period studied, lead-authorship of JCGBA articles is highly skewed towards men with only 34% women.
+A greater difference between men and women lead-authorship can be seen in CA, in 2009 there was no women lead-authors at all, and overall only 23% of the published papers of CA have women as lead-authors [@bardolph2018controlling].
+
 When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership they remain underrepresented in peer-reviewed journals.
@@
 In the University of California, Santa Barbara Gender Equity Project,@vanderwarker2018uscb report an imbalance in pay rates, for example, women make a greater proportion of people paid less \$60,000 and men outnumber women in salary brackets greater than \$60,000.
-(@vanderwarker2018uscb).
+[@vanderwarker2018uscb].
 
-<!-- summarise studies of citation practices that show a gap -->
-
-<!-- conference participation, check Bardolph 2018-->
-
-Something about Twitter
+Studies of authorship where that gender is inferred from the author's first name are common, but can give wrong results for non-English, androgynous, and uncommon names.
+@heathwrites avoided these limitations by conducting a survey that directly asked archaeologists for their self-identifications of gender, race/ethnicity and sexual orientation.
+In her intersectional study, @heathwrites a higher number of women publishing in archaeology than shown in previous studies, even in the studies of the same publications that she studied.
+@heathwrites explains this as a result of change over time, with her study including more recent publications than others, and concludes that most journals is slowly reaching gender parity.
 
@@
 For EAA we generated a total of `r EAA_stm_topic_number` topics, with `r EAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
-Although topics  significantly associated with men and women both include the word "Mediterranean", it is important to distinguish that females seems to be focusing on western Mediterranean while males seems to be focusing on northern Mediterranean Another  difference  is that males tends to have topics that are related to architecture, while the topic regarding research methods using isotopes seems to be more popular for females.
+Although topics significantly associated with men and women both include the word "Mediterranean", it is important to distinguish that females seems to be focusing on western Mediterranean while males seems to be focusing on northern Mediterranean.
+Another difference is that males tends to have topics that are related to architecture, while the topic regarding research methods using isotopes seems to be more popular for females.
 

### 2020-06-12 analysis/paper/paper.Rmd
abs

--- 2020-06-05 analysis/paper/paper.Rmd
@@
 
+# Abstract
+Disproportionate representation of genders has long existed among different academic fields.
+Rising interest in gender equality in society, generally, has stimulated concern about how much progress has been made towards gender equality in various areas in acamdemics.
+Analysis of peer-reviewed publication authorship shows that archaeology and other academic fields have long been dominated by men.
+We ask if gender disproportionality is evident in topic differences among archaeologists participating in major conferences, such as the Society of American Archaeology (SAA), European Association of Archaeologists (EAA) and the Computer Applications and Quantitative Methods in Archaeology (CAA) meetings.
+Do different gender of the participants in these archaeological conferences show topic difference between them?
+We analysed presenters' names in the published programs of archaeology conferences to infer gender.
+We found that there is some association between gender and topic.
+Being able to note the differences between topic of choice and gender participation is important to ensure equal participation in archaeology.
+We expect these findings to be similar in other fields especially with increased priority of an interest in gender equality in many research communities.
+
 # Introduction
@@
 When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field [@bardolph2014critical].
-<!--- something about increasing numbers of women's with PhD and increasing rates of participation -->
+According to American Anthropological Association's (AAA) _AnthroGuide_, there is a growing number of women participating in anthropology's graduate programs.
+For example, in the 2012-2013 school year, women consists of 64% of graduate students in anthropology programs and 65% of students that completed an anthropology doctoral degrees are women, whereas in 1972 32% of PhD are granted to women and 59% in 1995 [@bardolph2014critical].
 

### 2020-06-12 analysis/paper/paper.Rmd
expand results section

--- 2020-06-05 analysis/paper/paper.Rmd
@@
 In the SAA data we found a total of `r prettyNum(SAA18_row, big.mark = ",")` presentations.
-Of these we could identify the first-named author of `r count_gendered_SAA18` (`r perc_gendered_SAA18`%) presentations as either men (n=`r male_count_SAA18`) or women (n=`r female_count_SAA18`).
+Of these we could identify the first-named authors of `r count_gendered_SAA18` (`r perc_gendered_SAA18`%) presentations as either men (n=`r male_count_SAA18`) or women (n=`r female_count_SAA18`).
+In the SAA presentations we identified the optimum number of topics as `r SAA_stm_topic_number`, with `r SAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
+Eight of these topics were associated with men, and nine topics associated with women.
+One of the main difference that we can see between the topics of the genders are the locations that the presenters are interested in.
+Topics associated more with women include bioarchaeology, cemeteries, burials, shells, and ceramics.
+Topics associated more with men include survey and landscape archaeology, geoarchaeology, rituals, and regional studies.
+Women are more likely to work on locations such as Mesoamerica, Arizona and national parks, while men are more likely to have a topics about the Great Lakes, and Honduras.
+
+For the EAA data we have `r prettyNum(EAA18_row, big.mark = ","` presentations, and `r count_gendered_EAA18` (`r perc_gendered_EAA18`%) where the first author could be classified as either a man (n=`r male_count_EAA18`) or woman (n=`r female_count_EAA18`).
+For EAA we generated a total of `r EAA_stm_topic_number` topics, with `r EAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
+Seven of these topics were associated with men, and nine topics associated with women.
+Although topics significantly associated with men and women both include the word "Mediterranean", women seems to be focusing more on the western Mediterranean while men seems to be focusing on the northern Mediterranean.
+Topics associated more with women include burials, graves, museums, animals, diet, and ritual.
+Topics associated more with men include geoarchaeology, architecture, and towns.
+
 For the CAA meeting we have `r prettyNum(CAA18_row, big.mark = ",")` presentations, and `r count_gendered_CAA18` (`r perc_gendered_CAA18`%) first-named authors could be classified as either men (n=`r male_count_CAA18`) or women (n=`r female_count_CAA18`).
-For the EAA data we have `r prettyNum(EAA18_row, big.mark = ","` presentations, and `r count_gendered_EAA18` (`r perc_gendered_EAA18`%) where the first author could be classified as either a man (n=`r male_count_EAA18`) or woman (n=`r female_count_EAA18`).
+For CAA we generated a total of `r CAA_stm_topic_number` topics, with `r CAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
+Three of these topics were associated with men, and three topics associated with women.
+Topics associated more with women include archives, GIS, geophysics, collaboration, and teaching.
+Topics associated more with men include environmental and landscape archaeology.
+
+# Discussion
+
 This results in similar women-to-men ratios for the SAA (`r female_to_male_SAA18`) and EAA (`r female_to_male_EAA18`), but much lower for the CAA (`r female_to_male_CAA18`).
 
-In the SAA presentations we identified the optimum number of topics as `r SAA_stm_topic_number`, with `r SAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
-One of the main difference that we can see between the topics of the genders are the locations that the presenters are interested in.
-Women are more likely to have a topic at a specific region such as Mesoamerica, Arizona and national parks, while men are more likely to have a topic that are in the western regions.
-
-For EAA we generated a total of `r EAA_stm_topic_number` topics, with `r EAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
-Although topics significantly associated with men and women both include the word "Mediterranean", it is important to distinguish that females seems to be focusing on western Mediterranean while males seems to be focusing on northern Mediterranean.
-Another difference is that males tends to have topics that are related to architecture, while the topic regarding research methods using isotopes seems to be more popular for females.
-
-For CAA we generated a total of `r CAA_stm_topic_number` topics, with `r CAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
-In CAA's stm topic plot we can see a clearer pattern from the topics marked in red.
-For females from CAA, we see that topics that are associated with Geographic Information System (GIS) are more likely to be a topic for female presenters.
-Another female-associated topics from the CAA topic plot is topics that includes the word women, excav and inform.
-As for males from CAA, there doesn't seem to have a specific word that is more associated with males besides the word enviorn.
-
-# Discussion
+Compare topics between the three confs
 

### 2020-06-12 analysis/paper/paper.Rmd
editing background

@@
 
+# Abstract
+
+Disproportionate representation of genders has long existed among different academic fields.
+Rising interest in gender equality in society, generally, has stimulated concern about how much progress has been made towards gender equality in various areas in acamdemics.
+Analysis of peer-reviewed publication authorship shows that archaeology and other academic fields have long been dominated by men.
+We ask if gender disproportionality is evident in topic differences among archaeologists participating in major conferences, such as the Society of American Archaeology (SAA), European Association of Archaeologists (EAA) and the Computer Applications and Quantitative Methods in Archaeology (CAA) meetings.
+Do different gender of the participants in these archaeological conferences show topic difference between them?
+We analysed presenters' names in the published programs of archaeology conferences to infer gender.
+We found that there is some association between gender and topic.
+Being able to note the differences between topic of choice and gender participation is important to ensure equal participation in archaeology.
+We expect these findings to be similar in other fields especially with increased priority of an interest in gender equality in many research communities.
+
 # Introduction
@@
 
-When focusing on the field of archaeology, we can see that gender imbalance still occurs in archaeology and that seems to impact the kind of publications that are done in the field [@bardolph2014critical].
-<!--- something about increasing numbers of womens with PhD and increasing rates of participation -->
+According to the American Anthropological Association's (AAA) _AnthroGuide_, there is a growing number of women participating in anthropology's graduate programs.
+For example, in the 2012-2013 school year, women consists of 64% of graduate students in anthropology programs and 65% of students that completed an anthropology doctoral degrees are women, whereas in 1972 32% of PhD are granted to women and 59% in 1995 [@bardolph2014critical].
+Although women slightly outnumber men in anthropology graduate programs, previous work has shown substantial gender imbalances in publication practices in archaeology where men greatly outnumber women, especially as first authors in peer reviewed journal articles [@bardolph2014critical].
+One of the first studies of this imbalance is @bardolph2014critical, who looked at 7,381 members of the Society of American Archaeology and found 53% men and 47% women.
+She also looked at 1,601 articles published during 1990-2013 in five high-visibility archaeology journals and found a range of proportions of women to men authors of 0.32-0.73, with three of the five journals having a proportion of <0.5.
 
@@
 We are only able to infer binary male/female genders and assign the first names into these two categories.
-This has the unfortunate result of excluding other genders from the results, excluding them from our analysis.
+This has the unfortunate result of excluding or mis-identifying other genders from the results, excluding them from our analysis.
 We considered it impractical and invasive to write to each first-named presented to request their gender information.
@@
 It is important to note that the inferences of gender presented here are not self-identified by the presenters, but are computed probabilistically.
-Better quality and more representative data would result if presenters self-identified their gender, but currently these data are not available.
+Better quality and more representative data would result if presenters self-identified their gender to conference organisers, but currently these data are not available.
 We encourage conference organizers to collect gender data directly from participants to improve the representation of minority genders in future studies.
@@
 We identified the topic of each presentation by generating a structural topic model for all the presentations in each of the three conferences.
-Topic modeling is a machine learning method to automatically find related groups of words that resemble traditional 'themes' or 'topics' in a collection of documents.Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields [add citations].
+Topic modeling is a machine learning method to automatically find related groups of words that resemble traditional 'themes' or 'topics' in a collection of documents.
+Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields [add citations].
 This approach uses latent Dirichlet allocation to allow every word to be present in every topic, but with different weightings, such that the most heavily weighted 5-10 words of a topic often capture the essence of the topic as a coherent and familiar concept.
 When generating topic models, we have to first decide on the number of topics that the method will identify in our texts.
-The topic modeling algorithm then  assigns all words in every document into our set number of topics, and assigns each word a probability based on its per-topic-per-word probability, and each document (i.e.
+<!-- how did we do this?
+--> The topic modeling algorithm then assigns all words in every document into our set number of topics, and assigns each word a probability based on its per-topic-per-word probability, and each document (i.e.
 conference presentation title) is assigned a distribution of topics with different weights.
@@
 A structural topic model (STM) allows us to include, alongside the typical topic modeling process, other document-level metadata to analyse covariance between topics and metadata [@roberts2019package].
-Covariates of interest can be included into the prior distributions for document-topic proportions and
-topic-word distributions using a standard regression model (cite this one: http://scholar.harvard.edu/files/dtingley/files/topicmodelsopenendedexperiments.pdf).
+Covariates of interest can be included into the prior distributions for document-topic proportions and topic-word distributions using a standard regression model <!-- (cite this one: http://scholar.harvard.edu/files/dtingley/files/topicmodelsopenendedexperiments.pdf) -->.
 This means that we can examine the relationship between the topics and document-level variables of interest.

### 2020-06-13 analysis/paper/paper.Rmd
edits

@@
 Disproportionate representation of genders has long existed among different academic fields.
-Rising interest in gender equality in society, generally, has stimulated concern about how much progress has been made towards gender equality in various areas in acamdemics.
-Analysis of peer-reviewed publication authorship shows that archaeology and other academic fields have long been dominated by men.
-We ask if gender disproportionality is evident in topic differences among archaeologists participating in major conferences, such as the Society of American Archaeology (SAA), European Association of Archaeologists (EAA) and the Computer Applications and Quantitative Methods in Archaeology (CAA) meetings.
-Do different gender of the participants in these archaeological conferences show topic difference between them?
+Rising interest in gender equality in society, generally, has stimulated concern about how much progress has been made towards gender equality in various areas in academics.
+Analysis of authorship of peer-reviewed publications shows that archaeology and other academic fields have long been dominated by men.
+We ask if gender disproportionality is evident in the choice of topics that researcher present on among archaeologists participating in major conferences, such as the Society of American Archaeology (SAA), European Association of Archaeologists (EAA) and the Computer Applications and Quantitative Methods in Archaeology (CAA) meetings.
+Does the gender of the participants in these archaeological conferences correlate with the topics of their presentations?
 We analysed presenters' names in the published programs of archaeology conferences to infer gender.
-We found that there is some association between gender and topic.
-Being able to note the differences between topic of choice and gender participation is important to ensure equal participation in archaeology.
-We expect these findings to be similar in other fields especially with increased priority of an interest in gender equality in many research communities.
+We using machine learning to identify topics from presentation titles.
+We found that there is some association between gender and topics.
+Being aware of correlations between research topics and gender is important to ensure equitable participation in archaeology and unbiased training opportunities for students.
+We expect these findings to useful to instructors who prioritize gender equality in student and early career research activities.
 
@@
 
+how to use these results... implicit bias in assigment topics...
+
 # Conclusion

### 2020-06-19 analysis/paper/paper.Rmd
citations

@@
 When generating topic models, we have to first decide on the number of topics that the method will identify in our texts.
-<!-- how did we do this?
---> The topic modeling algorithm then assigns all words in every document into our set number of topics, and assigns each word a probability based on its per-topic-per-word probability, and each document (i.e.
+We have used the built in algorithm of the stm package based off of Lee and Mimno (2014) to find the number of topics [@mimno2014low].
+This method finds the number of topic by choosing anchor words and using statistical probability so that for each assumed topic a specific anchor word will only appear once, after this the algorithm will reconstruct the word co-occurance pattern for the non-anchor words to create a convex combination of the co-occurance patterns of those anchor words.
+After creating the convex the algorithm will then choose the farthest point from our subspace until the given number of anchors have been found.
+It is important to note that this method will not alway result in the same number of suggested topics, as it is based off of probability and dependent on the quality of the anchor words chosen.
+The topic modeling algorithm then assigns all words in every document into our set number of topics, and assigns each word a probability based on its per-topic-per-word probability, and each document (i.e.
 conference presentation title) is assigned a distribution of topics with different weights.
@@
 A structural topic model (STM) allows us to include, alongside the typical topic modeling process, other document-level metadata to analyse covariance between topics and metadata [@roberts2019package].
-Covariates of interest can be included into the prior distributions for document-topic proportions and topic-word distributions using a standard regression model <!-- (cite this one: http://scholar.harvard.edu/files/dtingley/files/topicmodelsopenendedexperiments.pdf) -->.
+Covariates of interest can be included into the prior distributions for document-topic proportions and topic-word distributions using a standard regression model [@roberts2014structural].
 This means that we can examine the relationship between the topics and document-level variables of interest.

### 2020-06-26 analysis/paper/paper.Rmd
citation

@@
 Topic modeling is a machine learning method to automatically find related groups of words that resemble traditional 'themes' or 'topics' in a collection of documents.
-Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields [add citations].
+Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields [@chang2009reading].
 This approach uses latent Dirichlet allocation to allow every word to be present in every topic, but with different weightings, such that the most heavily weighted 5-10 words of a topic often capture the essence of the topic as a coherent and familiar concept.

### 2020-06-26 analysis/paper/paper.Rmd
STM
update abstract for SAA proposal

--- 2020-06-19 analysis/paper/paper.Rmd
@@
 ---
-title: "Title Goes Here"
+title: "Correlations between gender and research topics at three major archaeology conferences"
 author:
@@
 
-Disproportionate representation of genders has long existed among different academic fields.
-Rising interest in gender equality in society, generally, has stimulated concern about how much progress has been made towards gender equality in various areas in academics.
-Analysis of authorship of peer-reviewed publications shows that archaeology and other academic fields have long been dominated by men.
-We ask if gender disproportionality is evident in the choice of topics that researcher present on among archaeologists participating in major conferences, such as the Society of American Archaeology (SAA), European Association of Archaeologists (EAA) and the Computer Applications and Quantitative Methods in Archaeology (CAA) meetings.
+Disproportionate representation of genders has long existed in many academic fields.
+Rising interest in gender equality in society generally has resulted in greater scrutiny on gender inequality in academic communities.
+Analysis of authorship of peer-reviewed publications shows that archaeology is similar to other academic fields in having long been dominated by men.
+We ask if gender disproportionality is evident in the choice of topics that archaeologists present on at major conferences, particularly the Society of American Archaeology (SAA), the European Association of Archaeologists (EAA) and the Computer Applications and Quantitative Methods in Archaeology (CAA) meetings.
 Does the gender of the participants in these archaeological conferences correlate with the topics of their presentations?
-We analysed presenters' names in the published programs of archaeology conferences to infer gender.
-We using machine learning to identify topics from presentation titles.
-We found that there is some association between gender and topics.
-Being aware of correlations between research topics and gender is important to ensure equitable participation in archaeology and unbiased training opportunities for students.
-We expect these findings to useful to instructors who prioritize gender equality in student and early career research activities.
+We analysed presenters' names in the published programs of these three archaeology conferences to infer gender.
+We then used machine learning to identify topics from presentation titles.
+We found that there are some associations between gender and topics.
+An awareness of these correlations between research topics and gender is important to ensure equitable participation in archaeology, and unbiased access to training opportunities for students.
+We expect these findings to be useful for instructors who prioritize gender equality in student and early career research activities.
 
@@
 Topic modeling is a machine learning method to automatically find related groups of words that resemble traditional 'themes' or 'topics' in a collection of documents.
-Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields [add citations].
+Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields <!-- [add citations] -->.
 This approach uses latent Dirichlet allocation to allow every word to be present in every topic, but with different weightings, such that the most heavily weighted 5-10 words of a topic often capture the essence of the topic as a coherent and familiar concept.
 When generating topic models, we have to first decide on the number of topics that the method will identify in our texts.
-We have used the built in algorithm of the stm package based off of Lee and Mimno (2014) to find the number of topics [@mimno2014low].
-This method finds the number of topic by choosing anchor words and using statistical probability so that for each assumed topic a specific anchor word will only appear once, after this the algorithm will reconstruct the word co-occurance pattern for the non-anchor words to create a convex combination of the co-occurance patterns of those anchor words.
+We have used the method proposed by @mimno2014low to find the number of topics.
+
+<!-- paraphrase this into one sentence : "The core idea of the spectral initialization is to approximately find the vertices of the convex hull of the word co-occurrences.
+The algorithm of Lee and Mimno (2014) projects the matrix into a low dimensional space using t-distributed stochastic neighbor embedding (Van Der Maaten 2014) and then exactly solves for the convex hull.
+This has the advantage of automatically selecting the number of topics.
+-->
+
+<!-- This method finds the optimum number of topics by choosing anchor words and using statistical probability so that for each assumed topic a specific anchor word will only appear once, after this the algorithm will reconstruct the word co-occurance pattern for the non-anchor words to create a convex combination of the co-occurance patterns of those anchor words.
 After creating the convex the algorithm will then choose the farthest point from our subspace until the given number of anchors have been found.
-It is important to note that this method will not alway result in the same number of suggested topics, as it is based off of probability and dependent on the quality of the anchor words chosen.
+It is important to note that this method will not aways result in the same number of suggested topics, as it is based off of probability and dependent on the quality of the anchor words chosen.
+-->
+
 The topic modeling algorithm then assigns all words in every document into our set number of topics, and assigns each word a probability based on its per-topic-per-word probability, and each document (i.e.

### 2020-07-10 analysis/paper/paper.Rmd
minor edits

@@
 
+In an article by @sinclair2016intellectual, he has compared the relative standings between men and women researchers in terms of number of citations made by other authors in the field of archaeology.
+Among the list of 50 most cited authors, only 6 of them (12%) are women, meaning that recognition in terms of citation are still mostly given to men, similar theme can be seen even if we increase to top 250 authors, with 19% being women [@sinclair2016intellectual].
+<!-- add highs and lows for women percentag from table 10 -->  One possible reason could be that men and women simply have preferences over different topics and this is one reason why we want to examine if there's some correlation between gender and topics of presentation.
+
 We have extended the study of ratios of authorships in the field of archaeology by specifically focusing on the gender imbalance that occurs in the topics that the authors present in major archaeological conferences.
@@
 In generating the topic model we excluded words that are common in English generally or in archaeology and so have little semantic value, for example, 'the', 'a', 'el', 'al', 'archaeology', etc.
+
+We identified the topic of each presentation by generating a structural topic model for all the presentations in each of the three conferences.
+Topic modeling is a machine learning method to automatically find related groups of words that resemble traditional 'themes' or 'topics' in a collection of documents.
+Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields [@chang2009reading].
+This approach uses latent Dirichlet allocation to allow every word to be present in every topic, but with different weightings, such that the most heavily weighted 5-10 words of a topic often capture the essence of the topic as a coherent and familiar concept.
+When generating topic models, we have to first decide on the number of topics that the method will identify in our texts.
+We have used the built in algorithm of the stm package based off of Lee and Mimno (2014) to find the number of topics [@mimno2014low].
+This method finds the number of topic by choosing anchor words and using statistical probability so that for each assumed topic a specific anchor word will only appear once, after this the algorithm will reconstruct the word co-occurance pattern for the non-anchor words to create a convex combination of the co-occurance patterns of those anchor words.
+After creating the convex the algorithm will then choose the farthest point from our subspace until the given number of anchors have been found.
+It is important to note that this method will not alway result in the same number of suggested topics, as it is based off of probability and dependent on the quality of the anchor words chosen.
+The topic modeling algorithm then assigns all words in every document into our set number of topics, and assigns each word a probability based on its per-topic-per-word probability, and each document (i.e.
+conference presentation title) is assigned a distribution of topics with different weights.
+To visualise the topic model output we use the 10 words with the highest weighting within each topic as keywords to represent each topic.
+In generating the topic model we excluded words that are common in English generally or in archaeology and so have little semantic value, for example, 'the', 'a', 'el', 'al', 'archaeology', etc.
+
+We have selected the number of topics using the algorithm by Lee and Mimno (2014).
+Using a random t-distribution neighbor embedding, with the goal of solving a convex hull of word co-occurrencecs by finding the vertices [@roberts2014stm].
 

### 2020-07-10 analysis/paper/paper.Rmd
resolve merge conflict

@@
 Topic modeling is a machine learning method to automatically find related groups of words that resemble traditional 'themes' or 'topics' in a collection of documents.
-Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields <!-- [add citations] -->.
-This approach uses latent Dirichlet allocation to allow every word to be present in every topic, but with different weightings, such that the most heavily weighted 5-10 words of a topic often capture the essence of the topic as a coherent and familiar concept.
-When generating topic models, we have to first decide on the number of topics that the method will identify in our texts.
-We have used the method proposed by @mimno2014low to find the number of topics.
-
-<!-- paraphrase this into one sentence : "The core idea of the spectral initialization is to approximately find the vertices of the convex hull of the word co-occurrences.
-The algorithm of Lee and Mimno (2014) projects the matrix into a low dimensional space using t-distributed stochastic neighbor embedding (Van Der Maaten 2014) and then exactly solves for the convex hull.
-This has the advantage of automatically selecting the number of topics.
--->
-
-<!-- This method finds the optimum number of topics by choosing anchor words and using statistical probability so that for each assumed topic a specific anchor word will only appear once, after this the algorithm will reconstruct the word co-occurance pattern for the non-anchor words to create a convex combination of the co-occurance patterns of those anchor words.
-After creating the convex the algorithm will then choose the farthest point from our subspace until the given number of anchors have been found.
-It is important to note that this method will not aways result in the same number of suggested topics, as it is based off of probability and dependent on the quality of the anchor words chosen.
--->
-
-The topic modeling algorithm then assigns all words in every document into our set number of topics, and assigns each word a probability based on its per-topic-per-word probability, and each document (i.e.
-conference presentation title) is assigned a distribution of topics with different weights.
-To visualise the topic model output we use the 10 words with the highest weighting within each topic as keywords to represent each topic.
-In generating the topic model we excluded words that are common in English generally or in archaeology and so have little semantic value, for example, 'the', 'a', 'el', 'al', 'archaeology', etc.
-
-We identified the topic of each presentation by generating a structural topic model for all the presentations in each of the three conferences.
-Topic modeling is a machine learning method to automatically find related groups of words that resemble traditional 'themes' or 'topics' in a collection of documents.
 Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields [@chang2009reading].
@@
 When generating topic models, we have to first decide on the number of topics that the method will identify in our texts.
+
 We have used the built in algorithm of the stm package based off of Lee and Mimno (2014) to find the number of topics [@mimno2014low].
@@
 In generating the topic model we excluded words that are common in English generally or in archaeology and so have little semantic value, for example, 'the', 'a', 'el', 'al', 'archaeology', etc.
-
-We have selected the number of topics using the algorithm by Lee and Mimno (2014).
-Using a random t-distribution neighbor embedding, with the goal of solving a convex hull of word co-occurrencecs by finding the vertices [@roberts2014stm].
 

### 2020-07-17 analysis/paper/paper.Rmd
table10

@@
 Among the list of 50 most cited authors, only 6 of them (12%) are women, meaning that recognition in terms of citation are still mostly given to men, similar theme can be seen even if we increase to top 250 authors, with 19% being women [@sinclair2016intellectual].
-<!-- add highs and lows for women percentag from table 10 -->  One possible reason could be that men and women simply have preferences over different topics and this is one reason why we want to examine if there's some correlation between gender and topics of presentation.
+After analyzing the most cited authors and their articles' thematic clusterings, it can be seen that there is relatively more women authors whose article is in the category of Archaeology of Early State Societies in the Americas (N and C America) with 46% women authors,the category of Archaeological chemistry (esp.
+lipids analysis) with 40% women authors, Isotope analysis with 30% women authors, and dating (esp.
+thermoluminescence and magnetics) with 30% women authors, while the average women author that are cited in the field of archaeology as a whole is around 14%.
+One possible reason could be that men and women simply have preferences over different topics and this is one reason why we want to examine if there's some correlation between gender and topics of presentation.
 

### 2020-07-18 analysis/paper/paper.Rmd
knits on my computer

--- 2020-07-10 analysis/paper/paper.Rmd
@@
 output: 
-    bookdown::html_document2:
+    bookdown::word_document2:
       fig_caption: yes
@@
 
+EAA18_stm <- 
+  stm_model_and_plot(conf_paper_titles = EAA18, 
+                     conf_gender_results = EAA18_gm, 
+                     first_name_appears_first = FALSE,
+                     max.em.its = 1000)
+
+saveRDS(EAA18_stm, here::here("analysis/data/derived-data/EAA18_topic_model_and_plot.rds"))
+
 CAA18_stm <-   
@@
 
-EAA18_stm <- 
-  stm_model_and_plot(conf_paper_titles = EAA18, 
-                     conf_gender_results = EAA18_gm, 
-                     first_name_appears_first = FALSE,
-                     max.em.its = 1000)
-
-saveRDS(EAA18_stm, here::here("analysis/data/derived-data/EAA18_topic_model_and_plot.rds"))
 ```
@@
 ```
+Overall we see similar women-to-men ratios for the SAA (`r female_to_male_SAA18`) and EAA (`r female_to_male_EAA18`), but much lower for the CAA (`r female_to_male_CAA18`).
+Even before we look into the gender preferences for specific topics within each conference, we already have a hint that computational topics may be less preferred by women than by men.
 
@@
 Topics associated more with men include survey and landscape archaeology, geoarchaeology, rituals, and regional studies.
-Women are more likely to work on locations such as Mesoamerica, Arizona and national parks, while men are more likely to have a topics about the Great Lakes, and Honduras.
+Women are more likely to work on locations such as Mesoamerica, Arizona and national parks, while men are more likely to present on topics about the Great Lakes, and Honduras.
 
-For the EAA data we have `r prettyNum(EAA18_row, big.mark = ","` presentations, and `r count_gendered_EAA18` (`r perc_gendered_EAA18`%) where the first author could be classified as either a man (n=`r male_count_EAA18`) or woman (n=`r female_count_EAA18`).
+For the EAA data we have `r prettyNum(EAA18_row, big.mark = ",")` presentations, and `r count_gendered_EAA18` (`r perc_gendered_EAA18`%) where the first author could be classified as either a man (n=`r male_count_EAA18`) or woman (n=`r female_count_EAA18`).
 For EAA we generated a total of `r EAA_stm_topic_number` topics, with `r EAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
 Seven of these topics were associated with men, and nine topics associated with women.
-Although topics significantly associated with men and women both include the word "Mediterranean", women seems to be focusing more on the western Mediterranean while men seems to be focusing on the northern Mediterranean.
+Although topics significantly associated with men and women both include the word "Mediterranean", women seem to be focusing more on the western Mediterranean while men seems to be focusing on the northern Mediterranean.
 Topics associated more with women include burials, graves, museums, animals, diet, and ritual.
@@
 
-This results in similar women-to-men ratios for the SAA (`r female_to_male_SAA18`) and EAA (`r female_to_male_EAA18`), but much lower for the CAA (`r female_to_male_CAA18`).
+Our results show correlations between the gender of the first-named presenter and the topics in their conference presentation.
+Although each of the three meetings have distinctive sets of topics, we can identify some common themes in the topics that correlate with
+
+Although the gender gap in academia has been shrinking in recent years, there remain substantial differences, which often result in underrepresentation of women in many fields and career stages, as well as disadvantages such as lower salaries and less access to high-prestige employment [@caprile2012meta].
+Even research into these gender biases suffers a bias, with @Cislak_Formanowicz_Saguy_2018 finding that articles on gender bias and race bias are funded less often and published in journals with a lower Impact Factor than articles on comparable instances of social discrimination.
 

### 2020-07-18 analysis/paper/function-for-structural-topic-model.R

@@ -181,7 +181,7 @@ ggplot() +
                colour = normal_colour) +
   geom_linerange(data = highlight_tbl,
                  aes(y = order,
-                     xmin =ci_lower,
+                     xmin = ci_lower,
                      xmax = ci_upper),
                  colour = "red") +
   geom_linerange(data = highlight_tbl,

### 2020-07-18 analysis/paper/functions-for-gender.R

@@ -97,6 +97,7 @@ d <-
   group_by(session) %>%
   distinct(`first-named speaker`, .keep_all = TRUE )
 
+
 # compute gender of first name
 library(gender)
 library(genderdata)
@@ -118,7 +119,8 @@ d_genders <-
 d_genders_unnest <-
   d_genders %>%
   mutate(speaker_gender = transpose(first_name_gender)[['result']]) %>%
-  unnest_legacy(speaker_gender) %>%
+  mutate(speaker_gender_chr = map(speaker_gender, ~mutate_all(.x, as.character))) %>%
+  unnest_legacy(speaker_gender_chr) %>%
   select(session, gender, name)
 
 return(genders_unnest = d_genders_unnest)

### 2020-07-18 analysis/paper/paper.Rmd
update dicsussion

@@
 
-Our results show correlations between the gender of the first-named presenter and the topics in their conference presentation.
-Although each of the three meetings have distinctive sets of topics, we can identify some common themes in the topics that correlate with
+Our results show significant correlations between the gender of the first-named presenter and the topics in their conference presentation in all three conferences.
+Although each of the three meetings have distinctive sets of topics, we can identify some common themes in the topics that correlate with gender.
 

### 2020-07-28 analysis/paper/paper.Rmd
add more discussion

@@
 Although each of the three meetings have distinctive sets of topics, we can identify some common themes in the topics that correlate with gender.
+We found that topics strongly associated with women at all three conferences relate to managing cultural heritage, GIS, and isotope analyses.
+Pair-wise shared topics between conferences also reveal the norms of each community.
+These contrasts and similarities reveal some of the choices that presenters make when deciding where to share their work.
+The SAA and EAA are defined mostly by the geographic region of their community, the Americas and Europe.
+They have more participants, and our model identified a higher number of topics in these two, compared to CAA.
+CAA, with its focus on computer applications and quantitative methods, has a much smaller community, and a small number of topics.
+So we might expect less in common between CAA and the two bigger conferences.
+At both the SAA and EAA we see that topics about burials, cemeteries, bodies, and graves are strongly associated with women.
+For the SAA and CAA the shared women-associated topics include learning and practice.
+For the EAA & CAA the common topics associated with women presenters are self-referential: women and gender.
+The absence of bioarchaeology and topics about human remains at CAA suggest that either archaeologists working on those topics do not recognise this conference as a meaningful place to present that work, or that there are some unrealised opportunities to apply computational and quantitative methods in bioarchaeological research.
+
+Topics that are strongly associated with men at all three conferences include geoarchaeology and geophysics.
+Topics associated with men at the SAA and CAA include built-location-based research, indicated by keywords such as house/town/site/fort.
+At both SAA and CAA we see landscape as a shared topic that is strongly associated with men.
+For the EAA and CAA there is little overlap in topics strongly associated with men.
+The geophysics topic is notable here because it is also in the CAA topic that is most strongly associated with women.
+It is strongly associated with men at the SAA and EAA, but also strongly associated with women at the CAA.
+This indicates that a topic is not immutably bound to either men or women, but that topics can shift in prominence among genders, depending on the context.
+This is important for understanding the relationship between genders, topics, and communities.
+These relationships are flexible and contingent, such that a topic-gender association is specific to a particular community (e.g.
+conference), and argues against topic-gender generalisations that attempt to transcend communities of practice.
+
+ideas vs things vs places
 
@@
 Even research into these gender biases suffers a bias, with @Cislak_Formanowicz_Saguy_2018 finding that articles on gender bias and race bias are funded less often and published in journals with a lower Impact Factor than articles on comparable instances of social discrimination.
-
-Compare topics between the three confs
 

### 2020-10-29 analysis/paper/paper.Rmd
add some discussion
edits to para about Sinclair

@@
 
-In an article by @sinclair2016intellectual, he has compared the relative standings between men and women researchers in terms of number of citations made by other authors in the field of archaeology.
-Among the list of 50 most cited authors, only 6 of them (12%) are women, meaning that recognition in terms of citation are still mostly given to men, similar theme can be seen even if we increase to top 250 authors, with 19% being women [@sinclair2016intellectual].
-After analyzing the most cited authors and their articles' thematic clusterings, it can be seen that there is relatively more women authors whose article is in the category of Archaeology of Early State Societies in the Americas (N and C America) with 46% women authors,the category of Archaeological chemistry (esp.
+@sinclair2016intellectual, compared the relative standings between men and women researchers in terms of number of citations made by other authors in the field of archaeology.
+Among the list of 50 most cited authors, only 6 of them (12%) are women, meaning that recognition by citation is still mostly given to men.
+A similar result can be seen if we increase the sample to the top 250 authors, with 19% women [@sinclair2016intellectual].
+After analyzing the most cited authors and their articles' thematic clusterings, we see that there are relatively more women authors in the category of Archaeology of Early State Societies in the Americas (N and C America) with 46% women authors.
+Other categories with high proportions of women are Archaeological chemistry (esp.
 lipids analysis) with 40% women authors, Isotope analysis with 30% women authors, and dating (esp.
-thermoluminescence and magnetics) with 30% women authors, while the average women author that are cited in the field of archaeology as a whole is around 14%.
-One possible reason could be that men and women simply have preferences over different topics and this is one reason why we want to examine if there's some correlation between gender and topics of presentation.
+thermoluminescence and magnetics) with 30% women authors.
+The average women author cited in the field of archaeology as a whole is around 14%.
+Sinclair concludes that the higher representation of women as cited authors on topics of laboratory analysis supports Gero's claim that women in archaeology were most active in laboratory-based activities rather than excavation/fieldwork related activities.
 
@@
 
+Talk about Thelwall... say that these gender differences exist, but
+
 ideas vs things vs places

### 2020-10-29 analysis/paper/paper.Rmd
some para in discussions

@@
 
-Talk about Thelwall... say that these gender differences exist, but
+Related work on gender differences in research topics has noted major differences in the proportions of men and women across many academic-related interests.
+One widely discussed theory, especially in psychology, to explain this is the theory of people-thing interest dimensions which proposes that men have stronger interest in things and their mechanisms, while women have a stronger interest in persons and their feelings [@su2009men].
+However, recent work by @thelwall2019gender shows that gender differences in choice of research topics cannot be fully explained by people-thing theory from psychology.
+Thewall et al.
+looked at 508,283 journal articles, classified the articles into research areas, and determined the authors' genders by matching names with the 1990 US census data.
+Instead of a people-thing contrast, they find that women more likely to use exploratory and qualitative methods rather than quantitative methods, and men show more interest in male in abstraction, and power/control fields.
+Similarly, ostapenko2018textual...
+In our data, we can similarly see in an interest from men in quantitative methods.
+However, our results do not so clearly divide the interests of the two genders we have data on, for example, with women are more prominent in GIS, and isotope analyses, which are highly quantitative areas.
 
-ideas vs things vs places
+Theories such as people-thing and Thewall et al's exploratory/abstraction assume that researchers freely make choices about their research preferences, and that these choices are influenced by inborn cognitive differences.
+However, these approaches do not fully explain gender differences in topic choices.
+A meta-analysis of gender and science research by the European Commission [@caprile2012meta] found that structural and life-course factors also profoundly shape the distribution of genders in research fields.
+Socialisation factors, such as media representation of scientists and family role models strongly shape career choices.
+Life-course factors include the 'rush hour' which is the time when family and academic demands collide and pivotal decisions are made about having children and establishing an academic career.
+Historically this is a time that leads to women making different choices to their male peers, but this can also affect caregivers of any gender.
+For example, caregivers may be less mobile to conduct field research or relocate for career advancement.
+These structural and life-course factors may add up to subtle, but pervasive exclusionary practices the skew gender representation in research fields that are independent of individual choices and preferences.
 

### 2020-11-12 analysis/paper/paper.Rmd
added mention of ostapenko2018textual

@@
 Instead of a people-thing contrast, they find that women more likely to use exploratory and qualitative methods rather than quantitative methods, and men show more interest in male in abstraction, and power/control fields.
-Similarly, ostapenko2018textual...
 In our data, we can similarly see in an interest from men in quantitative methods.
 However, our results do not so clearly divide the interests of the two genders we have data on, for example, with women are more prominent in GIS, and isotope analyses, which are highly quantitative areas.
+
+@ostapenko2018textual identified common topics in the personal statements of aspiring surgeons that were specific to men and women authors.
+Women tended to discuss surgery as working as a team, while men focused on their specific individual clinical experiences.
+@ostapenko2018textual propose that the differences between male and female statements may come from actual motivating factors for career goals and aspirations, or they may reflect differences in beliefs about what makes a successful personal statement.
+They could not evaluate which explanation was most important, but a key observation is that authors of the personal statements may not deliberately choose the themes for their statements.
+Instead, they may be using themes that they have received positive feedback on on the past, that has perhaps unconsciously biased them towards certain themes in their writing.
+Similarly archaeologists might choose topics for their research based on a pattern of positive or negative feedback on topic choices over time, and so their choices are not fully deliberate.
 

### 2021-03-15 analysis/paper/paper.Rmd
added basic plot of data

@@
 lynne_gender <-  gender("Lynne")
-
 ```
@@
 
-SAA18_gm <- readr::read_rds(here::here("analysis/data/derived-data/SAA18_gender_model.rds")) 
-EAA18_gm <- readr::read_rds(here::here("analysis/data/derived-data/EAA18_gender_model.rds"))
-CAA18_gm <- readr::read_rds(here::here("analysis/data/derived-data/CAA18_gender_model.rds"))
+SAA18_gm <- 
+  readr::read_rds(here::here("analysis/data/derived-data/SAA18_gender_model.rds"))
+
+EAA18_gm <- 
+  readr::read_rds(here::here("analysis/data/derived-data/EAA18_gender_model.rds"))
+
+CAA18_gm <- 
+  readr::read_rds(here::here("analysis/data/derived-data/CAA18_gender_model.rds"))
 
@@
 count_gendered_CAA18 <- nrow(CAA18_gender)
+
 perc_gendered_CAA18 <- round(nrow(CAA18_gender) / CAA18_row, 2) * 100
+
 female_count_CAA18 <- 
   length(which(CAA18_gender$gender == "female"))
+
 male_count_CAA18 <- 
   length(which(CAA18_gender$gender == "male"))
+
 female_to_male_CAA18 <- 
   round(female_count_CAA18 / male_count_CAA18, 1)
+
 missing_gender_CAA18 <-
@@
 ```
+
+```{r}
+# some basic visualisations to show the data
+list_confs <- 
+  list(SAA18_gender %>% 
+              mutate(session = as.character(session)),
+       CAA18_gender,
+       EAA18_gender %>% 
+              mutate(session = as.character(session))
+       )
+
+names(list_confs) <- c("SAA", "CAA", "EAA")
+
+all_conf_tbl <- 
+  bind_rows(list_confs,
+            .id = "conf")
+
+# how many papers per conf?
+ggplot(all_conf_tbl) +
+  aes(conf,
+      fill = gender) +
+  geom_bar(position = "dodge") +
+  theme_minimal(base_size = 16) +
+  theme(legend.position = c(.2,.60)) +
+  labs(x = "", 
+       y = "number of papers presented in 2018")
+```
+
 Overall we see similar women-to-men ratios for the SAA (`r female_to_male_SAA18`) and EAA (`r female_to_male_EAA18`), but much lower for the CAA (`r female_to_male_CAA18`).
@@
 
+A key limitation of this study is our concept of gender, and how we measure it.
+Like many social science studies, we have measured gender assuming it has two binary categories.
+However, we recognise that many cultures have long included more than two genders [@wilson1996we; @graham2004s], and there is a growing awareness of this in Western cultures also.
+A better measure of gender would represent it as a multifaceted spectrum [@tobin2010intrapsychics].
+The problem with our binary measure is that it does not reflect our current understanding of gender, forces people in misclassified categories, and is hostile to public acceptance and advocacy for transgender and nonbinary individuals.
+Our instrument for collecting gender data only returns two categories, so this has limited potential for overcoming this limitation.
+Instead, we recommend that conference organisers collect gender information directly from presenters using inclusive gender measures.
+For example, by providing nonbinary options for presenters on registration forms, and by asking about gender as an open‐ended question for participants to self-identify.
+We recommend that future work on gender in archaeology avoid 'othering' language when describing results.
+For example, reporting a sample as '150 participants (48% women; 49% men; 3% other)' violates ethical standards because such wording implies that binary gender is normal or appropriate, whereas trans and nonbinary gender is not (it is 'other').
+Our hope is that our study might be the last one to  discuss gender as a binary.
+
+# Conclusion
+
 Although the gender gap in academia has been shrinking in recent years, there remain substantial differences, which often result in underrepresentation of women in many fields and career stages, as well as disadvantages such as lower salaries and less access to high-prestige employment [@caprile2012meta].
@@
 how to use these results... implicit bias in assigment topics...
-
-# Conclusion
 

### 2022-02-07 analysis/paper/paper.Rmd
many updates on the text to prepare to send to the editors for feedback

@@
 Women are often under-represented for several reasons, including stereotyping that unfairly underestimates individuals' abilities, which often leads to lack of support for women in academic fields [@xu2008gender].
-In this paper we explore the relationship between gender and research topics to determine if women and men tend to focus more on certain topics than others.
+Although the gender gap in academia has been shrinking in recent years, there remain substantial differences, which often result in underrepresentation of women in many fields and career stages, as well as disadvantages such as lower salaries and less access to high-prestige employment [@caprile2012meta].
+Even research into these gender biases suffers a bias, with @Cislak_Formanowicz_Saguy_2018 finding that articles on gender bias and race bias are funded less often and published in journals with a lower Impact Factor than articles on comparable instances of social discrimination.
+In this paper we explore the relationship between gender and archaeological research topics to determine if women and men tend to focus more on certain topics than others.
 
-We used a computational method to identify topics from the titles of presentations delivered at meetings of the Society of American Archaeology (SAA, 2016-2019), the European Association of Archaeologists (EAA, 2018) and the Computer Applications and Quantitative Methods in Archaeology (CAA, 2017).
+We used a computational method to identify topics from the titles of presentations delivered at meetings of the Society of American Archaeology (SAA, 2018), the European Association of Archaeologists (EAA, 2018) and the Computer Applications and Quantitative Methods in Archaeology (CAA, 2018).
 We analysed the covariance of topics and genders of presenters to explore how gender ratios vary between these conferences and by topics.
@@
 
+Gender bias and imbalance has been a concern in archaeology since the earliest discussions of the social structure of the discipline [@Gero1983].
 According to the American Anthropological Association's (AAA) _AnthroGuide_, there is a growing number of women participating in anthropology's graduate programs.
-For example, in the 2012-2013 school year, women consists of 64% of graduate students in anthropology programs and 65% of students that completed an anthropology doctoral degrees are women, whereas in 1972 32% of PhD are granted to women and 59% in 1995 [@bardolph2014critical].
+For example, in the 2012-2013 school year, women consist of 64% of graduate students in anthropology programs and 65% of students that completed an anthropology doctoral degrees were women, whereas in 1972 32% of PhD are granted to women and 59% in 1995 [@bardolph2014critical].
 Although women slightly outnumber men in anthropology graduate programs, previous work has shown substantial gender imbalances in publication practices in archaeology where men greatly outnumber women, especially as first authors in peer reviewed journal articles [@bardolph2014critical].
-One of the first studies of this imbalance is @bardolph2014critical, who looked at 7,381 members of the Society of American Archaeology and found 53% men and 47% women.
+One of the first studies of this imbalance in publication practices is @bardolph2014critical, who looked at 7,381 members of the Society of American Archaeology and found 53% men and 47% women.
 She also looked at 1,601 articles published during 1990-2013 in five high-visibility archaeology journals and found a range of proportions of women to men authors of 0.32-0.73, with three of the five journals having a proportion of <0.5.
@@
 
-Studies of authorship where that gender is inferred from the author's first name are common, but can give wrong results for non-English, androgynous, and uncommon names.
+Studies of authorship where gender is inferred from the author's first name are common, but can give wrong results for non-English, androgynous, and uncommon names.
 @heathwrites avoided these limitations by conducting a survey that directly asked archaeologists for their self-identifications of gender, race/ethnicity and sexual orientation.
-In her intersectional study, @heathwrites a higher number of women publishing in archaeology than shown in previous studies, even in the studies of the same publications that she studied.
+In her intersectional study, @heathwrites found a higher number of women publishing in archaeology than shown in previous studies, even in the studies of the same publications that she studied.
 @heathwrites explains this as a result of change over time, with her study including more recent publications than others, and concludes that most journals is slowly reaching gender parity.
@@
 A similar result can be seen if we increase the sample to the top 250 authors, with 19% women [@sinclair2016intellectual].
-After analyzing the most cited authors and their articles' thematic clusterings, we see that there are relatively more women authors in the category of Archaeology of Early State Societies in the Americas (N and C America) with 46% women authors.
-Other categories with high proportions of women are Archaeological chemistry (esp.
+After analyzing the most cited authors and their articles' thematic clusterings, they found that there are relatively more women authors in the category of archaeology of early state societies in the Americas (N and C America) with 46% women authors.
+Other categories with high proportions of women are archaeological chemistry (esp.
 lipids analysis) with 40% women authors, Isotope analysis with 30% women authors, and dating (esp.
@@
 The average women author cited in the field of archaeology as a whole is around 14%.
-Sinclair concludes that the higher representation of women as cited authors on topics of laboratory analysis supports Gero's claim that women in archaeology were most active in laboratory-based activities rather than excavation/fieldwork related activities.
+Sinclair concludes that the higher representation of women as cited authors on topics of laboratory analysis supports Gero's [-@Gero1994] claim that women in archaeology were most active in laboratory-based activities rather than excavation/fieldwork related activities.
 
-We have extended the study of ratios of authorships in the field of archaeology by specifically focusing on the gender imbalance that occurs in the topics that the authors present in major archaeological conferences.
+In this paper we advance the study of gender and authorship in archaeology by specifically focusing on the gender imbalance that occurs in the topics that the authors present in major archaeological conferences.
 
@@
 
-We requested from conference organisers spreadsheets files of the publicly available program information for three major archaeology meetings held in 2018: the Society of American Archaeology (SAA, 2016-2019), the European Association of Archaeologists (EAA, 2018) and the Computer Applications and Quantitative Methods in Archaeology (CAA, 2017).
+We requested from conference organisers spreadsheets files of the publicly available program information for three major archaeology meetings held in 2018: the Society of American Archaeology (SAA), the European Association of Archaeologists (EAA) and the Computer Applications and Quantitative Methods in Archaeology (CAA).
 
-We estimated the gender of the first-named speaker for each presentation using the R programming language and the gender package [@lincolngender; @mihaljevic2019reflections; @blevins2015jane].
+We estimated the gender of the first-named speaker for each presentation using the R programming language and the 'gender' package [@lincolngender; @mihaljevic2019reflections; @blevins2015jane].
 We inferred the gender of personal names by looking each name up in the US Social Security Administration (SSA) baby name data, and calculating the overall probability that a given name was male or female [@mihaljevic2019reflections].
-If the proportion of people in the SSA data with a certain given name are recorded as female is 0.5 or higher, we predicted the gender of the presenter with this name as female.
+If the proportion of people in the SSA data with a certain given name are recorded as female is 0.5 or higher, we recorded the gender of the presenter with this name as female.
 For example, in the SSA dataset the the name "Lynne" returns `r lynne_gender$proportion_male` as proportion of individuals with this name who are recorded as male, and `r lynne_gender$proportion_female` as the proportion female, thus we inferred that people in our conference data who are called Lynne are women.
 
-This method of inferring gender has the advantage of speed, transparency and reproducibility, but also some substantial limitations that are important to be upfront about.
+This method of inferring gender has the advantage of rapidly determining gender for a large number of names in a transparent and reproducible way, but it also some substantial limitations that are important to be upfront about.
 We are only able to infer binary male/female genders and assign the first names into these two categories.
@@
 We encourage conference organizers to collect gender data directly from participants to improve the representation of minority genders in future studies.
-Our hope is that this work, despite its many limitations, will stimulate the collection of more reliable, justifiable, and useful data by conference organisers.
+Our hope is that this work, despite (or perhaps because of) its many limitations and biases, will stimulate the collection of more reliable, justifiable, and useful data by conference organisers.
 
 We identified the topic of each presentation by generating a structural topic model for all the presentations in each of the three conferences.
-Topic modeling is a machine learning method to automatically find related groups of words that resemble traditional 'themes' or 'topics' in a collection of documents.
+Topic modeling is a machine learning method that is widely used to automatically find related groups of words that resemble traditional 'themes' or 'topics' in large collections of documents.
 Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields [@chang2009reading].
-This approach uses latent Dirichlet allocation to allow every word to be present in every topic, but with different weightings, such that the most heavily weighted 5-10 words of a topic often capture the essence of the topic as a coherent and familiar concept.
-When generating topic models, we have to first decide on the number of topics that the method will identify in our texts.
+This approach uses latent Dirichlet allocation (LDA) to allow every word to be present in every topic, but with different weightings, such that the most heavily weighted 5-10 words of a topic often capture the essence of the topic as a coherent and familiar concept.
+When generating topic models, we have to first determine the number of topics that the method will identify in our texts.
+To prepare our data for topic modelling we excluded words that are common in English generally or in archaeology and so have little semantic value, for example, 'the', 'a', 'el', 'al', 'archaeology', etc.
+A list of these stopwords is included in our research compendium.
+We also applied a stemming procedure to words in our documents as part of our pre-processing.
+Stemming reduces similar words to their common root, for example, 'technological', 'technologies' and 'technology' all share the common stem of 'technolog'.
+Stemming is useful for removing repetition and transforming words to their base form to simplify interpretation of the output.
 
-We have used the built in algorithm of the stm package based off of Lee and Mimno (2014) to find the number of topics [@mimno2014low].
-This method finds the number of topic by choosing anchor words and using statistical probability so that for each assumed topic a specific anchor word will only appear once, after this the algorithm will reconstruct the word co-occurance pattern for the non-anchor words to create a convex combination of the co-occurance patterns of those anchor words.
-After creating the convex the algorithm will then choose the farthest point from our subspace until the given number of anchors have been found.
-It is important to note that this method will not alway result in the same number of suggested topics, as it is based off of probability and dependent on the quality of the anchor words chosen.
-The topic modeling algorithm then assigns all words in every document into our set number of topics, and assigns each word a probability based on its per-topic-per-word probability, and each document (i.e.
+We used the LDA algorithm implemented in the 'stm' package which uses the method of Lee and Mimno (2014) to find the optimum number of topics [@mimno2014low].
+This method finds the number of topics by choosing anchor words and using statistical probabilities so that for each assumed topic a specific anchor word will only appear once, after this the algorithm will reconstruct the word co-occurrence patterns for the non-anchor words to create a convex combination of the co-occurrence patterns of those anchor words.
+Then the algorithm will then choose the farthest point from our subspace until the given number of anchors have been found.
+It is important to note that because this method is probabilistic, it will not always result in the same number of suggested topics.
+The topic modeling algorithm then assigns all words in every document into our pre-determined number of topics, and assigns each word a probability based on its per-topic-per-word probability.
+Similarly, each document (i.e.
 conference presentation title) is assigned a distribution of topics with different weights.
 To visualise the topic model output we use the 10 words with the highest weighting within each topic as keywords to represent each topic.
-In generating the topic model we excluded words that are common in English generally or in archaeology and so have little semantic value, for example, 'the', 'a', 'el', 'al', 'archaeology', etc.
 
@@
 This means that we can examine the relationship between the topics and document-level variables of interest.
-In our case, we computed the relationship between the inferred the gender of the first-named presenter for each presentation, and the distribution of topics in each presentation.
-When we observed the topics for each presentation, we generated a regression where the topic is the outcome variable, and gender is the explanatory variable.
-This regression gives us insights into whether gender caused archaeologists to spend a larger portion of their presentation on a particular topic.
+In our case, we computed the relationship between the inferred the gender of the first-named presenter for each presentation, and the distribution of topics in that person's presentation.
+After we observed the topics for each presentation, we then generated a regression where the topic is the outcome variable, and gender is the explanatory variable.
+This regression gives us insights into whether there are non-random relationships between the inferred gender of an archaeologist and the topics of their presentation.
+Our data visualisations represent the estimated marginal effect (the x-axis values) of gender on each topic (distributed on the y-axis).
+Our visualisations show point estimates and 95% confidence intervals for each topic, and we interpret a statistically significant gender effect on the distribution of a topic when the 95% confidence interval does not includ also (also coloured red in our figures).
+
+# Reproducibility and open source materials {-}
+
+The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at xxx to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
+Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations.
+The raw data for our analysis are equivalent to the publicly available data available in the conference programs, but are organised here in a tabular structure convenient for computational analysis.
+All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
+The code is released under the MIT license, the data as CC-0, and figures as CC-BY, to enable maximum re-use.
 
@@
 They have more participants, and our model identified a higher number of topics in these two, compared to CAA.
-CAA, with its focus on computer applications and quantitative methods, has a much smaller community, and a small number of topics.
+CAA, with its focus on computer applications and quantitative methods, has a much smaller community, and a smaller number of topics.
 So we might expect less in common between CAA and the two bigger conferences.
@@
 For the EAA & CAA the common topics associated with women presenters are self-referential: women and gender.
-The absence of bioarchaeology and topics about human remains at CAA suggest that either archaeologists working on those topics do not recognise this conference as a meaningful place to present that work, or that there are some unrealised opportunities to apply computational and quantitative methods in bioarchaeological research.
+The absence of bioarchaeology and topics about human remains at CAA suggest that either archaeologists working on those topics do not recognise this conference as a meaningful place to present that work, and that there are some unrealised opportunities to apply computational and quantitative methods in bioarchaeological research.
 
@@
 Related work on gender differences in research topics has noted major differences in the proportions of men and women across many academic-related interests.
-One widely discussed theory, especially in psychology, to explain this is the theory of people-thing interest dimensions which proposes that men have stronger interest in things and their mechanisms, while women have a stronger interest in persons and their feelings [@su2009men].
+One early and widely discussed theory, especially in psychology, to explain this is the theory of people-thing interest dimensions which proposes that men have stronger interest in things and their mechanisms, while women have a stronger interest in people and their feelings [@Miner1922; @su2009men].
+@StoetGeary2022 used this theory in their investigation of gender differences in 473,260 adolescents’ aspirations to work in things-oriented (e.g., mechanic), people-oriented (e.g., nurse), and STEM (e.g., mathematician) careers across 80 countries and economic regions using data from the 2018 Programme for International Student Assessment (PISA).
+They found that boys generally aspired to a things-oriented or STEM occupation and  girls to a people-oriented occupation.
+Some of our results are consistent with this people-thing model, for example, 'people' topics such as burials, cemeteries, bodies, graves, collaboration, and teaching are associated more with women than men.
+
 However, recent work by @thelwall2019gender shows that gender differences in choice of research topics cannot be fully explained by people-thing theory from psychology.
@@
 Socialisation factors, such as media representation of scientists and family role models strongly shape career choices.
-Life-course factors include the 'rush hour' which is the time when family and academic demands collide and pivotal decisions are made about having children and establishing an academic career.
+Life-course factors include the 'rush hour' which is the time when family and academic demands collide and pivotal decisions are made about whether or not to have children and how much to invest or sacrifice for an academic career.
 Historically this is a time that leads to women making different choices to their male peers, but this can also affect caregivers of any gender.
 For example, caregivers may be less mobile to conduct field research or relocate for career advancement.
+This may explain the prominence of women in  GIS, and isotope analyses, which may be pursued without requiring travel for fieldwork.
 These structural and life-course factors may add up to subtle, but pervasive exclusionary practices the skew gender representation in research fields that are independent of individual choices and preferences.
 
-A key limitation of this study is our concept of gender, and how we measure it.
+A key limitation of this study is our concept of gender itself, and how we measure it.
 Like many social science studies, we have measured gender assuming it has two binary categories.
@@
 Our instrument for collecting gender data only returns two categories, so this has limited potential for overcoming this limitation.
-Instead, we recommend that conference organisers collect gender information directly from presenters using inclusive gender measures.
+To overcome this limitation, we recommend that conference organisers collect gender information directly from presenters using inclusive gender measures.
 For example, by providing nonbinary options for presenters on registration forms, and by asking about gender as an open‐ended question for participants to self-identify.
@@
 For example, reporting a sample as '150 participants (48% women; 49% men; 3% other)' violates ethical standards because such wording implies that binary gender is normal or appropriate, whereas trans and nonbinary gender is not (it is 'other').
-Our hope is that our study might be the last one to  discuss gender as a binary.
+Our hope is that our study might be the last one done by archaeologists to discuss gender as a binary.
 
@@
 
-Although the gender gap in academia has been shrinking in recent years, there remain substantial differences, which often result in underrepresentation of women in many fields and career stages, as well as disadvantages such as lower salaries and less access to high-prestige employment [@caprile2012meta].
-Even research into these gender biases suffers a bias, with @Cislak_Formanowicz_Saguy_2018 finding that articles on gender bias and race bias are funded less often and published in journals with a lower Impact Factor than articles on comparable instances of social discrimination.
+We find consistent associations of topics and genders across the three conferences.
+Like much previous work [@lippa2005gender], the associations we observed are not easily explained by conventional models such as people/thing or exploratory/abstraction models.
+We speculate that the patterns in our data may result from a combination subtle, implicit biases that shape the decisions made by early career researchers and their advisers and mentors, as a well as large scale structural constraints that make certain topics more accessible for one gender than others.
 
-how to use these results... implicit bias in assigment topics...
+Our results are important because they include scholars at the beginning of their career, such as students and post-doctoral researchers that have yet to publish in a peer reviewed journal, and thus not represented or underrepresented in studies of gender and publication patterns.
+As such, our findings have implications for how archaeologists are trained and mentored at the early stages of their career.
+First, instructors and mentors should be alert to unintentionally recommending students into classes, research projects and laboratory or field experiences that might appear to suit their gender, for example, 'people' topics for women students, rather than presenting students with a variety of topics and encouraging students to choose based on their interests.
+Second, senior archaeologists should be intentional about removing structural limitations to participation in certain topics by gender.
+For example, support for childcare expenses should be available for career-relevant travel, such as field work.
+This may help to equalize professional mobility for women researchers who may have limited capacity to travel with young children.
+Third, departments that train archaeologists should ensure that hiring practices support a high diversity of topics, and equal representation of genders to minimize historical effects of low diversity that will canalize gender-topic associations, making it harder for students to freely choose research topics.
+
+@Brown2018 found the low rate of women as professors, and other senior leadership positions in archaeology results in a confounding effect on female students' ability to find suitable models and mentors.
+Our work indicates that the negative effects of a low rate of women professors is not limited to female students, because without women professors, students of any gender may be unable to pursue certain research topics that currently tend to be pursued by women.
+Similarly, supporting womens' participation in fieldwork is likely to have a positive effect for junior researchers of all genders, who can benefit from increased sharing of expertise on topics currently concentrated among women researchers.
+It is our hope that studies such as ours will stimulate a more active consideration of gender-topic associations that will lead to a more equitable future for archaeology.
 
 # Acknowledgements
+
+Thanks to the organisers of the SAA, EAA and CAA meetings for sharing their data in a format suitable for quantitative analysis.
+Earlier versions of this research were presented University of Washington Undergraduate Research Symposium in 2020, and the SAA meetings in 2021.
+We're grateful for questions and feedback from those events that helped to improve this paper.
+Thanks to Catherine Jalbert and	Laura Heath-Stout for their encouragement and feedback.
 

### 2022-02-08 analysis/paper/paper.Rmd
add figure captions

@@
       email: yichun33@uw.edu
-      institute: [UofO]
+      institute: [UofW]
       correspondence: true
   - Ben Marwick:
-      email: fl@another.edu
-      institute: [UofA]
+      email: bmarwick@uw.edu
+      institute: [UofW]
       correspondence: false
 institute:
-  - UofO: University of One Place
-  - UofA: University of Another Place
+  - UofW: Department of Anthropology, University of Washington, Seattle
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
     bookdown::word_document2:
+      number_sections: no
       fig_caption: yes
@@
 csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
-abstract: |
-  Text of abstract
-keywords: |
-  keyword 1; keyword 2; keyword 3
-highlights: |
-  These are the highlights. 
 ---
@@
 Gender bias and imbalance has been a concern in archaeology since the earliest discussions of the social structure of the discipline [@Gero1983].
-According to the American Anthropological Association's (AAA) _AnthroGuide_, there is a growing number of women participating in anthropology's graduate programs.
+According to the American Anthropological Association's (AAA) *AnthroGuide*, there is a growing number of women participating in anthropology's graduate programs.
 For example, in the 2012-2013 school year, women consist of 64% of graduate students in anthropology programs and 65% of students that completed an anthropology doctoral degrees were women, whereas in 1972 32% of PhD are granted to women and 59% in 1995 [@bardolph2014critical].
@@
 One of the first studies of this imbalance in publication practices is @bardolph2014critical, who looked at 7,381 members of the Society of American Archaeology and found 53% men and 47% women.
-She also looked at 1,601 articles published during 1990-2013 in five high-visibility archaeology journals and found a range of proportions of women to men authors of 0.32-0.73, with three of the five journals having a proportion of <0.5.
+She also looked at 1,601 articles published during 1990-2013 in five high-visibility archaeology journals and found a range of proportions of women to men authors of 0.32-0.73, with three of the five journals having a proportion of \<0.5.
 
@@
 Another similar imbalance can be seen in terms of the authorship of articles or papers written.
-For example, for the authorship of the first authors of 1,104 articles published in the _Journal of Field Archaeology_ during 1974–2018, 72% of them are men [@heath2020guest].
+For example, for the authorship of the first authors of 1,104 articles published in the *Journal of Field Archaeology* during 1974--2018, 72% of them are men [@heath2020guest].
 @heath2020guest investigates the possibility that the gender gap in authorship is due to due to sexism in the peer review process.
-Out of 830 instances of peer review for where both the reviewer’s gender and the first author's gender could be determined, neither the first author’s gender, the reviewer’s gender, nor the combination of the two had a significant effect on the reviewer's recommendation.
+Out of 830 instances of peer review for where both the reviewer's gender and the first author's gender could be determined, neither the first author's gender, the reviewer's gender, nor the combination of the two had a significant effect on the reviewer's recommendation.
 
-@fulkerson2019dominates collected data on author's gender and occupational affiliation in peer reviewed journals (_American Antiquity_ (AQ), _Advances in Archaeological Practice_ (AAP), _Journal of California and Great Basin Anthropology_ (JCGBA), and _California Archaeology_ (CA)) and some non-peer-reviewed venues (the _SAA Archaeological Record_ (SAA Record) and the SCA _Proceedings_).
+@fulkerson2019dominates collected data on author's gender and occupational affiliation in peer reviewed journals (*American Antiquity* (AQ), *Advances in Archaeological Practice* (AAP), *Journal of California and Great Basin Anthropology* (JCGBA), and *California Archaeology* (CA)) and some non-peer-reviewed venues (the *SAA Archaeological Record* (SAA Record) and the SCA *Proceedings*).
 Among the 5,010 authors of 2,445 articles in their sample, 27.1% of first/single authors of peer-reviewed journal articles are women, and 72.9% are men.
@@
 When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership they remain underrepresented in peer-reviewed journals.
-@tushingham2017peer examined authorship gender trends in 1,599 papers in three journals, _Journal of California and Great Basin Anthropology/Journal of California Archaeology_ (JCGBA/JCA), _California Archaeology_ (CA) and a non-peer reviewed _Proceedings_ of the Society for California Archaeology (PSCA).
+@tushingham2017peer examined authorship gender trends in 1,599 papers in three journals, *Journal of California and Great Basin Anthropology/Journal of California Archaeology* (JCGBA/JCA), *California Archaeology* (CA) and a non-peer reviewed *Proceedings* of the Society for California Archaeology (PSCA).
 In a total of 2,617 authors, 844 (32.3%) were women, 1,762 (67.3%) were men, and 11 (0.4%) were gender unknown/ambiguous [@tushingham2017peer].
@@
 This similar trend of a gender gap in archaeology can also be seen in women's pay [@vanderwarker2018ucsb].
-In the University of California, Santa Barbara Gender Equity Project,@vanderwarker2018uscb report an imbalance in pay rates, for example, women make a greater proportion of people paid less \$60,000 and men outnumber women in salary brackets greater than \$60,000.
-[@vanderwarker2018uscb].
+In the University of California, Santa Barbara Gender Equity Project, @vanderwarker2018ucsb report an imbalance in pay rates, for example, women make a greater proportion of people paid less \$60,000 and men outnumber women in salary brackets greater than \$60,000.
 
@@
 library(here)
+library(tidyverse)
+library(ggplot2)
 # column names should be:
@@
 Our data visualisations represent the estimated marginal effect (the x-axis values) of gender on each topic (distributed on the y-axis).
-Our visualisations show point estimates and 95% confidence intervals for each topic, and we interpret a statistically significant gender effect on the distribution of a topic when the 95% confidence interval does not includ also (also coloured red in our figures).
+Our visualisations show point estimates and 95% confidence intervals for each topic, and we interpret a statistically significant gender effect on the distribution of a topic when the 95% confidence interval does not include zero.
+We use triangular points and red colouring to indicate the significant effects, and black circular points for non-significant effects.
+Our plots include summaries each topic represented by the top 12 stemmed words for each topic.
 
-# Reproducibility and open source materials {-}
+# Reproducibility and open source materials {.unnumbered}
 
@@
 
-```{r, eval=FALSE}
+```{r, eval=TRUE}
 
@@
 
-```
-
-```{r}
-# get plots
-library(tidyverse)
-
-# SAA -----
-SAA18_stm_plot <- 
-  readr::read_rds(here::here("analysis/data/derived-data/SAA18_topic_model_and_plot.rds"))
-
-SAA18_stm_plot$plot
-ggplot2::ggsave(here::here("analysis/figures/SAA18_stm_plot.png"), 
-       height = 7, 
-       width = 6)
-
-SAA_stm_topic_number <-
-  SAA18_stm_plot$model$settings$dim$K
-SAA_stm_topic_number_significant_covary <- 
-  nrow(layer_data(SAA18_stm_plot$plot, 3))
-
-# CAA -----
-CAA18_stm_plot <- 
-  readr::read_rds(here::here("analysis/data/derived-data/CAA18_topic_model_and_plot.rds"))
-
-CAA18_stm_plot$plot
-ggsave(here::here("analysis/figures/CAA18_stm_plot.png"), height = 7, width = 6)
-
-CAA_stm_topic_number <-
-  CAA18_stm_plot$model$settings$dim$K
-CAA_stm_topic_number_significant_covary <- 
-  nrow(layer_data(CAA18_stm_plot$plot, 3))
-
-# EAA -----
-EAA18_stm_plot <- 
-  readr::read_rds(here::here("analysis/data/derived-data/EAA18_topic_model_and_plot.rds"))
-
-EAA18_stm_plot$plot
-ggsave(here::here("analysis/figures/EAA18_stm_plot.png"), height = 7, width = 6)
-
-EAA_stm_topic_number <-
-  EAA18_stm_plot$model$settings$dim$K
-EAA_stm_topic_number_significant_covary <- 
-  nrow(layer_data(EAA18_stm_plot$plot, 3))
 ```
@@
 
-```{r}
+Overall we see similar women-to-men ratios for the SAA (`r female_to_male_SAA18`) and EAA (`r female_to_male_EAA18`), but much lower for the CAA (`r female_to_male_CAA18`) (Figure \@ref(fig:basic-viz)).
+Even before we look into the gender preferences for specific topics within each conference, we already have a hint that computational topics may be less preferred by women than by men.
+
+```{r basic-viz}
+#| fig.cap = "Frequencies of papers at each of the three conferences by gender"
+
 # some basic visualisations to show the data
@@
   bind_rows(list_confs,
-            .id = "conf")
+            .id = "conf") %>% 
+  mutate(gender = ifelse(gender == "female",
+                         "woman", "man"))
 
@@
   geom_bar(position = "dodge") +
-  theme_minimal(base_size = 16) +
+  # colours from https://blog.datawrapper.de/gendercolor/
+  scale_fill_discrete(type = c("#8625F5", "#1FC3AA")) +
+  theme_minimal(base_size = 14) +
   theme(legend.position = c(.2,.60)) +
@@
 
-Overall we see similar women-to-men ratios for the SAA (`r female_to_male_SAA18`) and EAA (`r female_to_male_EAA18`), but much lower for the CAA (`r female_to_male_CAA18`).
-Even before we look into the gender preferences for specific topics within each conference, we already have a hint that computational topics may be less preferred by women than by men.
+```{r saa-stm-plot}
+#| fig.cap = "Point estimates and 95% confidence intervals for the estimated 
+#| marginal effect of gender on each topic identified in titles of presentations at
+#| 2018 SAA meeting"
+
+# get plots
+library(tidyverse)
+
+# SAA -----
+SAA18_stm_plot <- 
+  readr::read_rds(here::here("analysis/data/derived-data/SAA18_topic_model_and_plot.rds"))
+
+SAA18_stm_plot$plot
+ggplot2::ggsave(here::here("analysis/figures/SAA18_stm_plot.png"), 
+       height = 7, 
+       width = 6)
+
+SAA_stm_topic_number <-
+  SAA18_stm_plot$model$settings$dim$K
+SAA_stm_topic_number_significant_covary <- 
+  nrow(ggplot2::layer_data(SAA18_stm_plot$plot, 3))
+```
 
@@
 Of these we could identify the first-named authors of `r count_gendered_SAA18` (`r perc_gendered_SAA18`%) presentations as either men (n=`r male_count_SAA18`) or women (n=`r female_count_SAA18`).
-In the SAA presentations we identified the optimum number of topics as `r SAA_stm_topic_number`, with `r SAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
+In the SAA presentations we identified the optimum number of topics as `r SAA_stm_topic_number`, with `r SAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter (Figure \@ref(fig:saa-stm-plot)).
 Eight of these topics were associated with men, and nine topics associated with women.
@@
 
+```{r eaa-stm-plot}
+#| fig.cap = "Point estimates and 95% confidence intervals for the estimated 
+#| marginal effect of gender on each topic identified in titles of presentations at
+#| 2018 EAA meeting"
+
+# EAA -----
+EAA18_stm_plot <- 
+  readr::read_rds(here::here("analysis/data/derived-data/EAA18_topic_model_and_plot.rds"))
+
+EAA18_stm_plot$plot
+ggsave(here::here("analysis/figures/EAA18_stm_plot.png"), height = 7, width = 6)
+
+EAA_stm_topic_number <-
+  EAA18_stm_plot$model$settings$dim$K
+EAA_stm_topic_number_significant_covary <- 
+  nrow(ggplot2::layer_data(EAA18_stm_plot$plot, 3))
+```
+
 For the EAA data we have `r prettyNum(EAA18_row, big.mark = ",")` presentations, and `r count_gendered_EAA18` (`r perc_gendered_EAA18`%) where the first author could be classified as either a man (n=`r male_count_EAA18`) or woman (n=`r female_count_EAA18`).
-For EAA we generated a total of `r EAA_stm_topic_number` topics, with `r EAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
+For EAA we generated a total of `r EAA_stm_topic_number` topics, with `r EAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter (Figure \@ref(fig:eaa-stm-plot)).
 Seven of these topics were associated with men, and nine topics associated with women.
@@
 
+```{r caa-stm-plot}
+#| fig.cap = "Point estimates and 95% confidence intervals for the estimated 
+#| marginal effect of gender on each topic identified in titles of presentations at
+#| 2018 CAA meeting"
+
+# CAA -----
+CAA18_stm_plot <- 
+  readr::read_rds(here::here("analysis/data/derived-data/CAA18_topic_model_and_plot.rds"))
+
+CAA18_stm_plot$plot
+ggsave(here::here("analysis/figures/CAA18_stm_plot.png"), height = 7, width = 6)
+
+CAA_stm_topic_number <-
+  CAA18_stm_plot$model$settings$dim$K
+CAA_stm_topic_number_significant_covary <- 
+  nrow(ggplot2::layer_data(CAA18_stm_plot$plot, 3))
+```
+
 For the CAA meeting we have `r prettyNum(CAA18_row, big.mark = ",")` presentations, and `r count_gendered_CAA18` (`r perc_gendered_CAA18`%) first-named authors could be classified as either men (n=`r male_count_CAA18`) or women (n=`r female_count_CAA18`).
-For CAA we generated a total of `r CAA_stm_topic_number` topics, with `r CAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter.
+For CAA we generated a total of `r CAA_stm_topic_number` topics, with `r CAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter (Figure \@ref(fig:eaa-stm-plot)).
 Three of these topics were associated with men, and three topics associated with women.
@@
 One early and widely discussed theory, especially in psychology, to explain this is the theory of people-thing interest dimensions which proposes that men have stronger interest in things and their mechanisms, while women have a stronger interest in people and their feelings [@Miner1922; @su2009men].
-@StoetGeary2022 used this theory in their investigation of gender differences in 473,260 adolescents’ aspirations to work in things-oriented (e.g., mechanic), people-oriented (e.g., nurse), and STEM (e.g., mathematician) careers across 80 countries and economic regions using data from the 2018 Programme for International Student Assessment (PISA).
+@StoetGeary2022 used this theory in their investigation of gender differences in 473,260 adolescents' aspirations to work in things-oriented (e.g., mechanic), people-oriented (e.g., nurse), and STEM (e.g., mathematician) careers across 80 countries and economic regions using data from the 2018 Programme for International Student Assessment (PISA).
 They found that boys generally aspired to a things-oriented or STEM occupation and girls to a people-oriented occupation.
@@
 A better measure of gender would represent it as a multifaceted spectrum [@tobin2010intrapsychics].
-The problem with our binary measure is that it does not reflect our current understanding of gender, forces people in misclassified categories, and is hostile to public acceptance and advocacy for transgender and nonbinary individuals.
+The problem with our binary measure is that it does not reflect our current understanding of gender, forces people in mis-classified categories, and is hostile to public acceptance and advocacy for transgender and nonbinary individuals.
 Our instrument for collecting gender data only returns two categories, so this has limited potential for overcoming this limitation.
@@
 <!-- The following line inserts a page break  -->
+
 \newpage
@@
 # References
+
 <!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
-<div id="refs"></div>
+
+::: {#refs}
+:::
 
@@
 
-The current Git commit details are:
-u
+The current Git commit details are: u
+
 ```{r}

### 2022-02-08 analysis/paper/function-for-structural-topic-model.R

@@ -210,7 +210,7 @@ ggplot() +
                 label = labels),
             size = highlight_size,
             colour = "red") +
-  xlab("        ← more men  ...  more women →") +
+  xlab("    ← more men  -  more women →") +
   ylab("Topic")
 
 # unlink(here::here("analysis/data/derived-data/the_model.rds"))

### 2022-02-08 analysis/paper/paper.Rmd
ready to send to editors

@@
 
-```{r, eval=TRUE}
+```{r, eval=FALSE}
 
@@
 #| marginal effect of gender on each topic identified in titles of presentations at 
-#| 2018 SAA meeting"
+#|  the 2018 SAA meeting"
 
@@
 #| marginal effect of gender on each topic identified in titles of presentations at 
-#| 2018 EAA meeting"
+#|  the 2018 EAA meeting"
 
@@
 #| marginal effect of gender on each topic identified in titles of presentations at 
-#| 2018 CAA meeting"
+#|  the 2018 CAA meeting"
 
@@
 For the CAA meeting we have `r prettyNum(CAA18_row, big.mark = ",")` presentations, and `r count_gendered_CAA18` (`r perc_gendered_CAA18`%) first-named authors could be classified as either men (n=`r male_count_CAA18`) or women (n=`r female_count_CAA18`).
-For CAA we generated a total of `r CAA_stm_topic_number` topics, with `r CAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter (Figure \@ref(fig:eaa-stm-plot)).
+For CAA we generated a total of `r CAA_stm_topic_number` topics, with `r CAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter (Figure \@ref(fig:caa-stm-plot)).
 Three of these topics were associated with men, and three topics associated with women.

### 2022-02-08 analysis/paper/paper.Rmd
add CI

@@
 
-The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at xxx to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
+The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at <https://github.com/yichun33/archyconfgender> to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
 Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations.

### 2022-02-08 README.Rmd
update binder to work with renv

--- 2020-03-02 README.Rmd
@@
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh///master?urlpath=rstudio)
+[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/yichun33/archyconfgender/master?urlpath=rstudio)
+[![.github/workflows/knit-in-Docker.yaml](https://github.com/yichun33/archyconfgender/actions/workflows/knit-in-Docker.yaml/badge.svg)](https://github.com/yichun33/archyconfgender/actions/workflows/knit-in-Docker.yaml)
 

### 2022-02-08 analysis/paper/function-for-structural-topic-model.R

@@ -13,6 +13,10 @@ stm_plot <- function(conf_paper_titles,
   library(readxl)
   library(quanteda)
   library(stm)
+  library(tm)
+  library(Rtsne)
+  library(geometry)
+  library(rsvd)
   library(dplyr)
 
   x <- conf_paper_titles
